---
title: "Project title"
author:
  - name: Author Name
    affiliation:
    - &aff1 Author Institution, City, Country.
    email: email@website.com
date: "2026-02-23"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
vignette: >
  %\VignetteIndexEntry{Project title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Programming environment

```{r Set random seed, include=FALSE}
seed <- 2026-02-23
```

```{r Load packages, include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(ggpubr)
library(pbapply)
library(broom)
library(brms)
library(cmdstanr)
library(broom.mixed)
library(mice)
filter <- dplyr::filter
cbind <- base::cbind
rbind <- base::rbind
```

```{r Load functions, include=FALSE}
lapply(list.files("R/", full.names = TRUE), source)
```

```{r Set theme, include=FALSE}
my_theme_set()
```

```{r Session information, echo=FALSE}
sessionInfo()
```

# Data Preprocessing

## Raw data

```{r Read code book, include=FALSE}
code_book <- read_csv("inst/extdata/code_book.csv", show_col_types = FALSE)
```

```{r Show code book, echo=FALSE}
show_table(code_book, "code_book", "Code book.")
```

```{r Column-type string, include=FALSE}
coltype_str <- paste0(code_book$col_type, collapse = "")
```

```{r Raw data, include=FALSE}
raw_data <- read_csv("inst/extdata/raw_data.csv", col_types = coltype_str)
```

```{r State row and column numbers of raw data, echo=FALSE}
cat("Rows: ", nrow(raw_data), "\n")
cat("Columns: ", ncol(raw_data))
```

## Select variable of interest

```{r Variable of interest, include=FALSE}
voi <- "pe"
```

```{r Rename the variable as outcome and put after identifier, include=FALSE}
proc_data <-
  raw_data |>
  rename(outcome = voi) |>
  select(id, outcome, everything())
```

# Data partition

```{r Determine partition sample size, include=FALSE}
whole_id <- proc_data$id
whole_n <- length(whole_id)
test_n <- ceiling(0.2 * whole_n)
dev_n <- whole_n - test_n
val_n <- ceiling(0.2 * dev_n)
train_n <- dev_n - val_n
```

```{r Randomly select identifiers for each partition, include=FALSE}
set.seed(seed)
test_id <- sample(whole_id, test_n)
dev_id <- setdiff(whole_id, test_id)

set.seed(seed)
val_id <- sample(dev_id, val_n)
train_id <- setdiff(dev_id, val_id)
```

```{r Create set data, include=FALSE}
set_data <-
  data.frame(id = train_id, set = "training") |>
  rbind(data.frame(id = val_id, set = "validation")) |>
  rbind(data.frame(id = test_id, set = "testing")) |>
  mutate(set = factor(set, c("training", "validation", "testing")))
```

```{r Join set data to outcome data, include=FALSE}
outcome_set_data <-
  proc_data |>
  select(id, outcome) |>
  left_join(set_data, by = "id")
```

```{r Summarize outcome class per set, include=FALSE}
if(class(outcome_set_data$outcome) == "numeric"){
  outcome_set_summary <-
    outcome_set_data |>
    group_by(set) |>
    summarize(n = n(), .groups = "drop")
  
  outcome_set_plot <-
    outcome_set_summary |>
    ggplot(aes(set, n)) +
    geom_col()
}else{
  outcome_set_summary <-
    outcome_set_data |>
    group_by(set, outcome) |>
    summarize(n = n(), .groups = "drop")
  
  outcome_set_plot <-
    outcome_set_summary |>
    ggplot(aes(set, n, fill = outcome)) +
    geom_col()
}
```

```{r Show summary per set, echo=FALSE, fig.height=90/25.4, fig.width=190/25.4}
outcome_set_plot
```

```{r Caption outcome class summary per set, echo=FALSE}
show_figure_legend("outcome_set_plot", "Data partition.")
```

```{r Split samples using the identifiers for train set, include=FALSE}
train_data <- slice(proc_data, match(train_id, id))
```

# QQ plot

```{r Select numerical variables, include=FALSE}
num_data <- select_if(train_data, is.numeric)
```

```{r State number of QQ plots, echo=FALSE}
cat("Number of QQ plots (= # of numeric variables): ", ncol(num_data))
```

```{r Create custom function for QQ plot}
qq_plotter <- function(variable, data){
  data |>
    `colnames<-`(str_replace_all(colnames(data), variable, "variable")) |>
    ggplot(aes(sample = variable)) +
  	geom_qq() +
  	geom_qq_line() +
    xlab("Normal +/- SD") +
    ylab(variable) +
    theme(axis.title = element_text(size = 10))
}
```

```{r Create QQ plots, include=FALSE}
if(ncol(num_data) > 0){
  qq_plots <- pblapply(colnames(num_data), qq_plotter, num_data)
  
  qq_ncol <- 4
  qq_nrow <- ceiling(length(qq_plots) / qq_ncol)
  qq_plots <- ggarrange(plotlist = qq_plots, ncol = qq_ncol, nrow = qq_nrow)
}
```

```{r Show QQ plots, echo=FALSE, fig.height=270/25.4, fig.width=190/25.4}
if(ncol(num_data) > 0){
  qq_plots
}
```

```{r Caption QQ plots, echo=FALSE}
if(ncol(num_data) > 0){
  show_figure_legend("qq_plots", "QQ plots.")
}
```

# Data transformation

```{r Variables for squared-root test-transformation, include=FALSE}
ttrans_var_sqrt <- c("plgf", "sflt1_plgf")
```

```{r Variables for squared test-transformation, include=FALSE}
ttrans_var_sq <- c()
```

```{r Variables for log test-transformation, include=FALSE}
ttrans_var_log <- c()
```

```{r State test-ransformed variables, echo=FALSE}
ttrans_var <- c(ttrans_var_sqrt, ttrans_var_sq, ttrans_var_log)

cat(
  paste0("Test-transformed variables (", length(ttrans_var), "): ")
  , ttrans_var
)
```

```{r Create custom function for squared transformation}
sq <- function(x){
  x^2
}
```

```{r Data test-transformation using numerical train set}
if(ncol(num_data) > 0 & length(ttrans_var) > 0){
  num_data_ttrans <-
    num_data |>
    mutate_at(ttrans_var_sqrt, sqrt) |>
    mutate_at(ttrans_var_sq, sq) |>
    mutate_at(ttrans_var_log, log)
}
```

```{r Create QQ test-trans, include=FALSE}
if(ncol(num_data_ttrans) > 0 & length(ttrans_var) > 0){
  qt_plots <- pblapply(ttrans_var, qq_plotter, num_data_ttrans)
  
  qt_ncol <- 4
  qt_nrow <- ceiling(length(qt_plots) / qt_ncol)
  qt_plots <- ggarrange(plotlist = qt_plots, ncol = qt_ncol, nrow = qt_nrow)
}
```

```{r Show QQ test-trans, echo=FALSE, fig.height=54/25.4, fig.width=190/25.4}
if(ncol(num_data_ttrans) > 0 & length(ttrans_var) > 0){
  qt_plots
}
```

```{r Caption QQ test-trans, echo=FALSE}
if(ncol(num_data_ttrans) > 0 & length(ttrans_var) > 0){
  show_figure_legend("qt_plots", "QQ plots after data transformation.")
}
```

```{r Variables for squared-root transformation, include=FALSE}
trans_var_sqrt <- c("sflt1_plgf")
```

```{r Variables for squared transformation, include=FALSE}
trans_var_sq <- c()
```

```{r Variables for log transformation, include=FALSE}
trans_var_log <- c()
```

```{r State transformed variables, echo=FALSE}
trans_var <- c(trans_var_sqrt, trans_var_sq, trans_var_log)
cat(paste0("Transformed variables (", length(trans_var), "): "), trans_var)
```

```{r State categorized numerical variables, echo=FALSE}
cat_num_var <- setdiff(ttrans_var, trans_var)
cat_num_var <- c(cat_num_var, "uta_bn", "parity")
cat(
  paste0("Categorized numerical variables (", length(cat_num_var), "): ")
  , cat_num_var
)
```

```{r Create custom function categorize column using pre-defined cutoff}
categorize_from_table <- function(data, cutoff_table){
  
  if(nrow(cutoff_table) == 0) return(data)
  
  cols <- unique(cutoff_table$column)
  
  for(col_name in cols){
    
    rules <-
      cutoff_table |>
      filter(column == col_name) |>
      arrange(cutoff)
    
    breaks <- c(rules$cutoff, Inf)
    labels <- rules$category
    
    data[[col_name]] <-
      cut(
        data[[col_name]]
        , breaks = breaks
        , labels = labels
        , right = FALSE
        , include.lowest = TRUE
      )
  }
  
  mutate_at(data, cols, factor)
}
```

```{r Read pre-defined cutoff for categorization, include=FALSE}
category_cutoff <-
  read_csv("inst/extdata/category_cutoff.csv", show_col_types = FALSE)
```

```{r Transform data for all sets, include=FALSE}
if(ncol(num_data) > 0){
  
  proc_data_trans <- proc_data
  
  # Numerical data transformation
  if(length(trans_var) > 0){
    proc_data_trans <-
      proc_data_trans |>
      mutate_at(trans_var_sqrt, sqrt) |>
      mutate_at(trans_var_sq, sq) |>
      mutate_at(trans_var_log, log)
  }
  
  # Numerical data categorization
  code_book_trans <- code_book
  if(length(intersect(cat_num_var, unique(category_cutoff$column))) > 0){
    proc_data_trans <-
      proc_data_trans |>
      categorize_from_table(category_cutoff)
    
    code_book_trans <-
      read_csv("inst/extdata/code_book_trans.csv", show_col_types = FALSE)
  }
}
```

```{r Obtain train set from transformed dataset, include=FALSE}
train_data_trans <- slice(proc_data_trans, match(train_id, id))
```

# Determine Tukey's fences

```{r Select numerical variables from transformed dataset, include=FALSE}
num_data_trans <- select_if(train_data_trans, is.numeric)
```

```{r State number of numeric variables after transformation, echo=FALSE}
cat("Number of numeric variables after transformation: ", ncol(num_data_trans))
```

```{r Create custom function to determine Tukey fences}
tukey_fence <- function(x, which = c("lb", "ub"), na.rm = TRUE){
	which <- match.arg(which)

	q1 <- quantile(x, 0.25, na.rm = na.rm)
	q3 <- quantile(x, 0.75, na.rm = na.rm)
	iqr <- q3 - q1

	if(which == "lb"){
		q1 - 1.5 * iqr
	}else{
		q3 + 1.5 * iqr
	}
}
```

```{r Determine Tukey fences from training set, include=FALSE}
if(ncol(num_data_trans) > 0){
  tukey_fence_lb <-
    num_data_trans |>
    summarize_all(tukey_fence, which = "lb") |>
    gather(col_name, tukey_lb)
  
  tukey_fence_ub <-
    num_data_trans |>
    summarize_all(tukey_fence, which = "ub") |>
    gather(col_name, tukey_ub)
  
  tukey_fences <-
    tukey_fence_lb |>
    left_join(tukey_fence_ub, by = join_by(col_name))
}
```

# Outliers handling 

```{r Flag outliers using Tukey fences for all sets, include=FALSE}
outliers <-
  proc_data_trans |>
  select(colnames(num_data_trans)) |>
  mutate(seq = seq(n())) |>
  gather(col_name, value, -seq) |>
  left_join(tukey_fences, by = join_by(col_name)) |>
  mutate(is_outlier = value < tukey_lb | value > tukey_ub) |>
  select(seq, col_name, is_outlier) |>
  spread(col_name, is_outlier) |>
  arrange(seq) |>
  select(colnames(num_data_trans))
```

```{r Summarize outliers for train set, include=FALSE}
outliers_sum <-
  select(proc_data_trans, id) |>
  cbind(outliers) |>
  slice(match(train_id, id)) |>
  select(-id) |>
  summarize_all(mean, na.rm = TRUE) |>
  gather(variable, p) |>
  mutate(p = round(p * 100, 2))
```

```{r Show outliers summary, echo=FALSE}
show_table(outliers_sum, "outliers_sum", "Outliers summary in train set.")
```

```{r Remove outliers by assigning values to missing, include=FALSE}
num_proc_data_outlier_rm <-
  proc_data_trans |>
  select(colnames(num_data_trans))

num_proc_data_outlier_rm[as.matrix(outliers)] <- NA
```

```{r Bind back outlier-removed data for all sets, include=FALSE}
proc_data_outlier_rm <-
  proc_data_trans |>
  select(setdiff(colnames(proc_data_trans), colnames(num_data_trans))) |>
  cbind(num_proc_data_outlier_rm) |>
  select(colnames(proc_data_trans))
```

```{r Obtain train set from outlier-removed dataset, include=FALSE}
train_data_outlier_rm <- slice(proc_data_outlier_rm, match(train_id, id))
```

```{r Add missingness indicators, include=FALSE}
train_data_ms <-
  train_data_outlier_rm |>
  select_if(\(x) any(is.na(x))) |>
  mutate_all(is.na) |>
  mutate_all(as.numeric) |>
  mutate_all(factor) |>
  rename_all(paste0, "_ms")

train_data_ms <-
  train_data_outlier_rm |>
  cbind(train_data_ms)
```

# Normality assessment and test

```{r Assess normality by QQ plot, include=FALSE}
normal_qq <- c("age", "height")
```

```{r State normal numerical variables by QQ plots, echo=FALSE}
cat("Normal numerical variables by QQ plot: ",normal_qq)
```

```{r List non-normal variables by QQ plot, include=FALSE}
non_normal_qq <- setdiff(colnames(num_data_trans), normal_qq)
```

```{r Conduct Shapiro-Wilk normality test, include=FALSE}
sw_results <-
  num_data_trans |>
  select_at(non_normal_qq) |>
  lapply(shapiro.test) |>
  lapply(tidy) |>
  imap(~ cbind(variable = .y, .x)) |>
  reduce(rbind) |>
  select(-method)
```

```{r Show Shapiro-Wilk normality test, echo=FALSE}
show_table(sw_results, "sw_results", "Shapiro-Wilk normality test results.")
```

```{r List normal variables by Shapiro-Wilk normality test, include=FALSE}
normal_sw <-
  sw_results |>
  filter(p.value > 0.05) |>
  pull(variable)
```

```{r List normal variables by Shapiro-Wilk normality test, include=FALSE}
non_normal_sw <- setdiff(non_normal_qq, normal_sw)
```

```{r Obtain normality for numerical columns, include=FALSE}
normality_data <-
  data.frame(col_name = colnames(num_data_trans)) |>
  mutate(normal = col_name %in% c(normal_qq, normal_sw))
```

# Perfect/partial separation

```{r Obtain scale for each column name except identifier, include=FALSE}
scale_data <-
    data.frame(
     scale =
       train_data_ms |>
       select(-id) |>
       sapply(class)
   ) |>
  rownames_to_column(var = "col_name")
```

```{r Identify all possible pair combinations, include=FALSE}
comb_pair <-
  combn(scale_data$col_name, 2) |>
  t() |>
  as.data.frame() |>
  left_join(
    rename(scale_data, V1 = col_name, scale_V1 = scale)
    , by = join_by(V1)
  ) |>
  left_join(
    rename(scale_data, V2 = col_name, scale_V2 = scale)
    , by = join_by(V2)
  )
```

```{r Create custom function to summarize variable pairs}
pair_summary <- function(data, V1, V2){
  result <-
    data |>
    group_by_at(c(V1, V2)) |>
    summarize(n = n(), .groups = "drop") |>
    `colnames<-`(c("value_V1", "value_V2", "n"))
  
  expand.grid(value_V1 = levels(data[[V1]]), value_V2 = levels(data[[V2]])) |>
    left_join(result, by = join_by(value_V1, value_V2)) |>
    mutate(n = ifelse(is.na(n), 0, n)) |>
    mutate(V1 = !!V1, V2 = !!V2)
}
```

```{r Perfect or partial separation data for categorical pairs, include=FALSE}
ps_data <-
  comb_pair |>
  filter(scale_V1 == "factor" & scale_V2 == "factor") |>
  select(V1, V2) |>
  pmap(\(V1, V2) pair_summary(train_data_ms, V1, V2)) |>
  reduce(rbind) |>
  group_by(V1, V2) |>
  summarize(ps = sum(n == 0) > 0, .groups = "drop")
```

# Correlation matrix

```{r Obtain number of categories for categorical variables, include=FALSE}
level_data <-
  train_data_ms |>
  select_if(is.factor) |>
  summarize_all(\(x) length(intersect(levels(x), unique(x)))) |>
  gather(col_name, level_n)
```

```{r Identify untestable pairs, include=FALSE}
comb_pair_untestable <-
  comb_pair |>
  filter(str_remove_all(V1, "_ms$") == str_remove_all(V2, "_ms$"))
```

```{r Join combinations with normality-separation-level data, include=FALSE}
comb_pair_normality_ps_level <-
  comb_pair |>
  anti_join(comb_pair_untestable, by = join_by(V1, V2, scale_V1, scale_V2)) |>
  left_join(
    rename(normality_data, V1 = col_name, normal_V1 = normal)
    , by = join_by(V1)
  ) |>
  left_join(
    rename(normality_data, V2 = col_name, normal_V2 = normal)
    , by = join_by(V2)
  ) |>
  left_join(ps_data, by = join_by(V1, V2))  |>
  left_join(
    rename(level_data, V1 = col_name, level_V1 = level_n)
    , by = join_by(V1)
  ) |>
  left_join(
    rename(level_data, V2 = col_name, level_V2 = level_n)
    , by = join_by(V2)
  ) |>
  mutate(
    normal_V1 = ifelse(is.na(normal_V1), TRUE, normal_V1)
    , normal_V2 = ifelse(is.na(normal_V2), TRUE, normal_V2)
    , ps = ifelse(is.na(ps), FALSE, ps)
    , level_V1 = ifelse(is.na(level_V1), 0, level_V1)
    , level_V2 = ifelse(is.na(level_V2), 0, level_V2)
  )
```

## Numerical-to-numerical pairs

```{r Identify numerical-to-numerical pairs, include=FALSE}
comb_pair_num_num <-
  comb_pair_normality_ps_level |>
  filter(scale_V1 == "numeric" & scale_V2 == "numeric")
```

```{r Identify normal-to-normal pairs, include=FALSE}
comb_pair_normal_normal <-
  comb_pair_num_num |>
  filter(normal_V1 & normal_V2) |>
  select(V1, V2)
```

```{r Identify non-normal-to-normal pairs, include=FALSE}
comb_pair_non_normal_normal <-
  comb_pair_num_num |>
  select(V1, V2) |>
  anti_join(comb_pair_normal_normal, join_by(V1, V2))
```

```{r Create custom function for correlation test}
correlation_test <- function(data, V1, V2, method){
  cor.test(train_data_ms[[V1]], train_data_ms[[V2]], method = "pearson") |>
    tidy() |>
    mutate(V1 = !!V1, V2 = !!V2) |>
    select(V1, V2, everything())
}
```

```{r Conduct Pearson correlation tests, include=FALSE}
if(nrow(comb_pair_normal_normal) > 0){
  cm_pearson <-
    comb_pair_normal_normal |>
    pmap(\(V1, V2) correlation_test(train_data_ms, V1, V2, "pearson")) |>
    reduce(rbind)
}
```

```{r Conduct Spearman correlation tests, include=FALSE}
if(nrow(comb_pair_non_normal_normal) > 0){
  cm_spearman <-
    comb_pair_non_normal_normal |>
    pmap(\(V1, V2) correlation_test(train_data_ms, V1, V2, "spearman")) |>
    reduce(rbind)
}
```

## Categorical-to-categorical pairs

```{r Identify categorical-to-categorical pairs, include=FALSE}
comb_pair_cat_cat <-
  comb_pair_normality_ps_level |>
  filter(scale_V1 == "factor" & scale_V2 == "factor")
```

```{r Identify pairs with separation, include=FALSE}
comb_pair_ps <-
  comb_pair_cat_cat |>
  filter(ps)
```

```{r Identify separation pairs with >2-category V1, include=FALSE}
comb_pair_ps_gt2_cat_V1 <-
  comb_pair_ps |>
  filter(level_V1 > 2) |>
  select(V1, V2)
```

```{r Create custom function for Bayesian categorical regression}
bayesian_cat_reg_fast <- function(data, V1, V2){
  data2 <-
    data |>
    select(all_of(c(V1, V2))) |>
    drop_na() |>
    droplevels()

  form <- reformulate(termlabels = V2, response = V1)

  result<-
    brm(
      formula = form
      , family = categorical()
      , data = data2
      , backend = "cmdstanr"
      , chains = 1
      , iter = 500
      , warmup = 250
      , refresh = 0
      , silent = 2
      , algorithm = "pathfinder"
    )
  
  tidy(result) |>
    mutate(V1 = !!V1, V2 = !!V2) |>
    select(V1, V2, everything())
}
```

```{r Conduct Bayesian categorical regression, include=FALSE}
if(nrow(comb_pair_ps_gt2_cat_V1) > 0){
  cm_bayesian_cat_reg <-
    comb_pair_ps_gt2_cat_V1 |>
    pmap(\(V1, V2) bayesian_cat_reg_fast(train_data_ms, V1, V2)) |>
    reduce(rbind)
}
```

```{r Identify separation pairs with 2-category V1, include=FALSE}
comb_pair_ps_2_cat_V1 <-
  comb_pair_ps |>
  filter(level_V1 == 2) |>
  select(V1, V2)
```

```{r Create custom function for Bayesian Bernoulli logistic regression}
bayesian_bern_reg_fast <- function(data, V1, V2){
  data2 <-
    data |>
    select(all_of(c(V1, V2))) |>
    drop_na() |>
    droplevels()

  form <- reformulate(termlabels = V2, response = V1)

  result <-
    brm(
      formula = form
      , family = bernoulli()
      , data = data2
      , backend = "cmdstanr"
      , chains = 1
      , iter = 500
      , warmup = 250
      , refresh = 0
      , silent = 2
      , algorithm = "pathfinder"
    )
  
  tidy(result) |>
    mutate(V1 = !!V1, V2 = !!V2) |>
    select(V1, V2, everything())
}
```

```{r Conduct Bayesian Bernoulli logistic regression, include=FALSE}
if(nrow(comb_pair_ps_2_cat_V1) > 0){
  cm_bayesian_bern_reg <-
    comb_pair_ps_2_cat_V1 |>
    pmap(\(V1, V2) bayesian_bern_reg_fast(train_data_ms, V1, V2)) |>
    reduce(rbind)
}
```

```{r Identify pairs without separation, include=FALSE}
comb_pair_non_ps <-
  comb_pair_cat_cat |>
  filter(!ps) |>
  select(V1, V2)
```

```{r Create custom function for Fisher exact test}
fisher_exact_test_fast <- function(data, V1, V2){
  result <-
    fisher.test(
      data[[V1]], data[[V2]]
      , workspace = 1e7, simulate.p.value = TRUE
    )
  
  tidy(result) |>
    select(p.value, method, alternative) |>
    mutate(V1 = !!V1, V2 = !!V2) |>
    select(V1, V2, everything())
}
```

```{r Conduct Fisher exact test, include=FALSE}
if(nrow(comb_pair_non_ps) > 0){
  cm_fisher <-
    comb_pair_non_ps |>
    pmap(\(V1, V2) fisher_exact_test_fast(train_data_ms, V1, V2)) |>
    reduce(rbind)
}
```

## Mixed-scale pairs

```{r Identify categorical-to-numerical pairs, include=FALSE}
comb_pair_cat_num <-
  comb_pair_normality_ps_level |>
  filter(scale_V1 == "factor" & scale_V2 == "numeric")
```

```{r Identify >2-categorical-to-norm-numerical pairs, include=FALSE}
comb_pair_cat_gt2_num_norm <-
  comb_pair_cat_num |>
  filter(level_V1 > 2) |>
  filter(normal_V2) |>
  select(V1, V2) |>
  mutate(inv = FALSE)
```

```{r Identify >2-categorical-to-non-norm-numerical pairs, include=FALSE}
comb_pair_cat_gt2_num_non_norm <-
  comb_pair_cat_num |>
  filter(level_V1 > 2) |>
  filter(!normal_V2) |>
  select(V1, V2) |>
  mutate(inv = FALSE)
```

```{r Identify 2-categorical-to-norm-numerical pairs, include=FALSE}
comb_pair_cat_2_num_norm <-
  comb_pair_cat_num |>
  filter(level_V1 == 2) |>
  filter(normal_V2) |>
  select(V1, V2) |>
  mutate(inv = FALSE)
```

```{r Identify 2-categorical-to-non-norm-numerical pairs, include=FALSE}
comb_pair_cat_2_num_non_norm <-
  comb_pair_cat_num |>
  filter(level_V1 == 2) |>
  filter(!normal_V2) |>
  select(V1, V2) |>
  mutate(inv = FALSE)
```

```{r Identify numerical-to-categorical pairs, include=FALSE}
comb_pair_num_cat <-
  comb_pair_normality_ps_level |>
  filter(scale_V1 == "numeric" & scale_V2 == "factor")
```

```{r Identify norm-numerical-to->2-categorical pairs, include=FALSE}
comb_pair_num_norm_cat_gt2 <-
  comb_pair_num_cat |>
  filter(level_V2 > 2)  |>
  filter(normal_V1) |>
  select(V1 = V2, V2 = V1) |>
  mutate(inv = TRUE)
```

```{r Identify non-norm-numerical-to->2-categorical pairs, include=FALSE}
comb_pair_num_non_norm_cat_gt2 <-
  comb_pair_num_cat |>
  filter(level_V2 > 2)  |>
  filter(!normal_V1) |>
  select(V1 = V2, V2 = V1) |>
  mutate(inv = TRUE)
```

```{r Identify norm-numerical-to-2-categorical pairs, include=FALSE}
comb_pair_num_norm_cat_2 <-
  comb_pair_num_cat |>
  filter(level_V2 == 2)  |>
  filter(normal_V1) |>
  select(V1 = V2, V2 = V1) |>
  mutate(inv = TRUE)
```

```{r Identify non-norm-numerical-to-2-categorical pairs, include=FALSE}
comb_pair_num_non_norm_cat_2 <-
  comb_pair_num_cat |>
  filter(level_V2 == 2)  |>
  filter(!normal_V1) |>
  select(V1 = V2, V2 = V1) |>
  mutate(inv = TRUE)
```

```{r Join >2-categorical and norm-numerical pairs, include=FALSE}
comb_pair_mixed_gt2_norm <-
  rbind(comb_pair_cat_gt2_num_norm, comb_pair_num_norm_cat_gt2)
```

```{r Create custom function for one-way ANOVA test}
one_way_anova_test <- function(data, V1, V2, inv){
  data <- data.frame(V1 = data[[V1]], V2 = data[[V2]])
  
  result <- aov(V2 ~ V1, data = data)
  
  result <-
    tidy(result) |>
    mutate(V1 = !!V1, V2 = !!V2)
  
  if(inv){
    result |>
      mutate(term = ifelse(term == "V1", "V2", term)) |>
      select(V1 = V2, V2 = V1, everything())
  }else{
    result |>
      select(V1, V2, everything())
  }
}
```

```{r Conduct one-way ANOVA test, include=FALSE}
if(nrow(comb_pair_mixed_gt2_norm) > 0){
  cm_one_way_anova <-
    comb_pair_mixed_gt2_norm |>
    pmap(\(V1, V2, inv) one_way_anova_test(train_data_ms, V1, V2, inv)) |>
    reduce(rbind)
}
```

```{r Join >2-categorical and non-norm-numerical pairs, include=FALSE}
comb_pair_mixed_gt2_non_norm <-
  rbind(comb_pair_cat_gt2_num_non_norm, comb_pair_num_non_norm_cat_gt2)
```

```{r Create custom function for Kruskal-Wallis test}
kruskal_wallis_test <- function(data, V1, V2, inv){
  V2_group_V1 <-
    data[[V1]] |>
    unique() |>
    lapply(\(x) data[[V2]][data[[V1]] == x])
  
  result <- kruskal.test(V2_group_V1)
  
  result <-
    tidy(result) |>
    mutate(V1 = !!V1, V2 = !!V2)
  
  if(inv){
    result |>
      select(V1 = V2, V2 = V1, everything())
  }else{
    result |>
      select(V1, V2, everything())
  }
}
```

```{r Conduct Kruskal-Wallis test, include=FALSE}
if(nrow(comb_pair_mixed_gt2_non_norm) > 0){
  cm_kruskal_wallis <-
    comb_pair_mixed_gt2_non_norm |>
    pmap(\(V1, V2, inv) kruskal_wallis_test(train_data_ms, V1, V2, inv)) |>
    reduce(rbind)
}
```

```{r Join 2-categorical and norm-numerical pairs, include=FALSE}
comb_pair_mixed_2_norm <-
  rbind(comb_pair_cat_2_num_norm, comb_pair_num_norm_cat_2)
```

```{r Create custom function for Welch two-sample t-test}
welch_two_sample_t_test <- function(data, V1, V2, inv){
  V2_group_V1_1 = data[[V2]][data[[V1]] == unique(data[[V1]])[1]]
  V2_group_V1_2 = data[[V2]][data[[V1]] == unique(data[[V1]])[2]]
  
  result <-
    try(
      t.test(
        V2_group_V1_1
        , V2_group_V1_2
        , alternative = "two.sided"
        , paired = FALSE
        , var.equal = FALSE
      )
    )
  
  if(is.character(result)){
    result <-
      data.frame(
        estimate = as.numeric(NA)
        , estimate1 = as.numeric(NA)
        , estimate2 = as.numeric(NA)
        , statistic = as.numeric(NA)
        , p.value = as.numeric(NA)
        , parameter = as.numeric(NA)
        , conf.low = as.numeric(NA)
        , conf.high = as.numeric(NA)
        , method =
          paste0(
            "Welch Two Sample t-test ("
            , paste0(result, collapse = ". ")
            , ")"
          )
        , alternative = "two.sided"
      )
  }else{
    result <- tidy(result)
  }
  
  result <- mutate(result, V1 = !!V1, V2 = !!V2)
  
  if(inv){
    result |>
      select(V1 = V2, V2 = V1, everything())
  }else{
    result |>
      select(V1, V2, everything())
  }
}
```

```{r Conduct Welch two-sample t-test, include=FALSE}
if(nrow(comb_pair_mixed_2_norm) > 0){
  cm_welch_two_sample_t <-
    comb_pair_mixed_2_norm |>
    pmap(\(V1, V2, inv) welch_two_sample_t_test(train_data_ms, V1, V2, inv)) |>
    reduce(rbind)
}
```

```{r Join 2-categorical and non-norm-numerical pairs, include=FALSE}
comb_pair_mixed_2_non_norm <-
  rbind(comb_pair_cat_2_num_non_norm, comb_pair_num_non_norm_cat_2)
```

```{r Create custom function for Wilcoxon rank-sum test}
wilcoxon_rank_sum_test <- function(data, V1, V2, inv){
  V2_group_V1_1 = data[[V2]][data[[V1]] == unique(data[[V1]])[1]]
  V2_group_V1_2 = data[[V2]][data[[V1]] == unique(data[[V1]])[2]]
  
  result <-
    try(
      wilcox.test(
        V2_group_V1_1
        , V2_group_V1_2
        , alternative = "two.sided"
        , paired = FALSE
      )
    )
  
  if(is.character(result)){
    result <-
      data.frame(
        statistic = as.numeric(NA)
        , p.value = as.numeric(NA)
        , method =
          paste0(
            "Welch Two Sample t-test ("
            , paste0(result, collapse = ". ")
            , ")"
          )
        , alternative = "two.sided"
      )
  }else{
    result <- tidy(result)
  }
  
  result <- mutate(result, V1 = !!V1, V2 = !!V2)
  
  if(inv){
    result |>
      select(V1 = V2, V2 = V1, everything())
  }else{
    result |>
      select(V1, V2, everything())
  }
}
```

```{r Conduct Welch two-sample t-test, include=FALSE}
if(nrow(comb_pair_mixed_2_non_norm) > 0){
  cm_wilcoxon_rank_sum <-
    comb_pair_mixed_2_non_norm |>
    pmap(\(V1, V2, inv) wilcoxon_rank_sum_test(train_data_ms, V1, V2, inv)) |>
    reduce(rbind)
}
```

## Multiple testing correction

```{r Create custom function to translate CI to p-value}
translate_ci_to_p_value <- function(data, V1, V2){
  data |>
    filter(V1 == !!V1 & V2 == !!V2) |>
    filter(!str_detect(term, "\\(Intercept\\)")) |>
    filter(!(conf.low <= 0 & conf.high >= 0)) |>
    summarize(n_sig = n()) |>
    mutate(p.value = ifelse(n_sig > 0, 0, 1)) |>
    filter(!is.na(p.value)) |>
    mutate(V1 = V1, V2 = V2) |>
    select(V1, V2, p.value)
}
```

```{r Translate CI to p-value for Bayesian categorical regression, include=FALSE}
cm_bayesian_cat_reg_p_value <-
  cm_bayesian_cat_reg |>
  select(V1, V2) |>
  unique() |>
  pmap(\(V1, V2) translate_ci_to_p_value(cm_bayesian_cat_reg, V1, V2)) |>
  reduce(rbind) |>
  mutate(method = "Bayesian categorical regression")
```

```{r Translate CI to p-value for Bayesian Bernoulli log reg, include=FALSE}
cm_bayesian_bern_reg_p_value <-
  cm_bayesian_bern_reg |>
  select(V1, V2) |>
  unique() |>
  pmap(\(V1, V2) translate_ci_to_p_value(cm_bayesian_bern_reg, V1, V2)) |>
  reduce(rbind) |>
  mutate(method = "Bayesian Bernoulli logistic regression")
```

```{r Combine results from all correlation tests, include=FALSE}
correlation_results <- list()

if(nrow(comb_pair_untestable) > 0){
  correlation_results <-
    correlation_results |>
    c(list(
        untestable =
          comb_pair_untestable |>
          select(V1, V2) |>
          mutate(p.value = as.numeric(NA), method = "None (untestable)")
      )
    )
}

if(nrow(comb_pair_normal_normal) > 0){
  correlation_results <- c(correlation_results, list(pearson = cm_pearson))
}

if(nrow(comb_pair_non_normal_normal) > 0){
  correlation_results <- c(correlation_results, list(spearman = cm_spearman))
}

if(nrow(comb_pair_ps_gt2_cat_V1) > 0){
  correlation_results <-
    c(correlation_results, list(bayesian_cat_reg = cm_bayesian_cat_reg_p_value))
}

if(nrow(comb_pair_ps_2_cat_V1) > 0){
  correlation_results <-
    correlation_results |>
    c(list(bayesian_bern_reg = cm_bayesian_bern_reg_p_value))
}

if(nrow(comb_pair_non_ps) > 0){
  correlation_results <- c(correlation_results, list(fisher = cm_fisher))
}

if(nrow(comb_pair_mixed_gt2_norm) > 0){
  correlation_results <-
    correlation_results |>
    c(list(
        one_way_anova =
          cm_one_way_anova |>
          filter(term != "Residuals") |>
          mutate(method = "One-way ANOVA")
      )
    )
}

if(nrow(comb_pair_mixed_gt2_non_norm) > 0){
  correlation_results <-
    c(correlation_results, list(kruskal_wallis = cm_kruskal_wallis))
}

if(nrow(comb_pair_mixed_2_norm) > 0){
  correlation_results <-
    c(correlation_results, list(welch_two_sample_t = cm_welch_two_sample_t))
}

if(nrow(comb_pair_mixed_2_non_norm) > 0){
  correlation_results <-
    c(correlation_results, list(wilcoxon_rank_sum = cm_wilcoxon_rank_sum))
}
```

```{r Combine p-values from all correlation tests, include=FALSE}
correlation_p_values <-
  correlation_results |>
  lapply(select, V1, V2, p.value, method) |>
  reduce(rbind)
```

```{r Compute FDRs from p-values by Benjamini-Hochberg method, include=FALSE}
correlation_fdr <-
  correlation_p_values |>
  mutate(
    FDR =
      p.value |>
      p.adjust(method = "BH")
    , sig =
      ifelse(is.na(FDR), "Not tested", ifelse(FDR <= 0.05, "Yes", "No")) |>
      factor()
    , method_no =
      as.numeric(factor(method, unique(method))) |>
      factor()
  ) |>
  select(sig, FDR, everything())
```

## Visualize correlation by significance

```{r Create custom function to plot correlation matrix}
cor_matrix_plotter <- function(data, triangle = "lower"){
  vars <- union(as.character(data$V1), as.character(data$V2))

  data_plot <-
    data |>
    mutate(
      V1 = factor(V1, levels = vars),
      V2 = factor(V2, levels = rev(vars)),
      sig = factor(sig, levels = c("No", "Yes", "Not tested"))
    ) |>
    mutate(
      x_id = as.integer(factor(V1, levels = vars)),
      y_var = as.character(V2),
      y_id = match(y_var, rev(vars))
    )

  if(triangle == "lower"){
    data_plot <- data_plot |> filter(x_id + y_id <= length(vars) + 1)
  }else if(triangle == "upper"){
    data_plot <- data_plot |> filter(x_id + y_id >= length(vars) + 1)
  }

  ggplot(data_plot, aes(x = V1, y = V2, fill = sig)) +
    geom_tile(color = "white") +
    geom_text(aes(label = method_no), size = 3) +
    scale_fill_manual(
      values = c(
        "No" = "grey80",
        "Yes" = "steelblue",
        "Not tested" = "grey40"
      )
    ) +
    coord_equal() +
    theme_minimal() +
    theme(
      panel.grid = element_blank(),
      axis.title = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
    labs(fill = "Significant")
}
```

```{r Plot cor matrix, include=FALSE}
correlation_matrix <-
  comb_pair |>
  select(V1, V2) |>
  left_join(correlation_fdr, by = join_by(V1, V2)) |>
  cor_matrix_plotter()
```

```{r Show corr matrix, echo=FALSE, fig.height=150/25.4, fig.width=190/25.4}
correlation_matrix
```

```{r Caption corr matrix, echo=FALSE}
show_figure_legend("correlation_matrix", "Correlation matrix.")
```

```{r Obtain correlation method numbers, include=FALSE}
cm_method_no <-
  correlation_matrix$data |>
  select(no = method_no, method) |>
  unique() |>
  arrange(no)
```

```{r Show correlation method numbers, echo=FALSE}
show_table(cm_method_no, "cm_method_no", "Correlation methods.")
```

# Missing value imputation

```{r Create custom function to make MICE predictor matrix, include=FALSE}
make_mice_pred_matrix <- function(
  corr_data
  , data_names = NULL
  , sig_value = "Yes"
  , ms_suffix = "_ms"
){
  # Variables only, exclude missingness indicators
  if(is.null(data_names)){
    data_names <-
      union(
        as.character(corr_data$V1)
        , as.character(corr_data$V2)
      )

    data_names <-
      data_names[!grepl(paste0(ms_suffix, "$"), data_names)] |>
      sort()
  }

  pred_matrix <-
    matrix(
      0
      , nrow = length(data_names)
      , ncol = length(data_names)
      , dimnames = list(data_names, data_names)
    )

  corr_yes <-
    corr_data |>
    filter(sig == sig_value) |>
    select(V1, V2) |>
    distinct()

  for(v in data_names){
    ms_v <- paste0(v, ms_suffix)

    preds_from_v <-
      corr_yes |>
      filter(V1 == v) |>
      pull(V2) |>
      as.character()

    preds_from_ms <-
      corr_yes |>
      filter(V1 == ms_v) |>
      pull(V2) |>
      as.character()

    preds <-
      c(preds_from_v, preds_from_ms) |>
      unique()

    preds <-
      preds[
        preds %in% data_names &
        preds != v
      ]

    pred_matrix[v, preds] <- 1
  }

  pred_matrix
}
```

```{r Create MICE predictor matrix}
pred_matrix <-
  correlation_matrix$data |>
  make_mice_pred_matrix(colnames(select(train_data_outlier_rm, -id)))
```

```{r Fit and save MICE model, include=FALSE}
imp_results <-
  mice(
    data = select(train_data_outlier_rm, -id)
    , method = 'pmm'
    , m = 10
    , seed = seed
    , predictorMatrix = pred_matrix
    , print = FALSE
  )

saveRDS(imp_results, "data/imp_results.rds")
```

```{r Impute missing values for all set, include=FALSE}
proc_data_imputed <-
  imp_results |>
  mice.mids(newdata = select(proc_data_outlier_rm, -id))|>
  complete(1)

proc_data_imputed <-
  cbind(select(proc_data_outlier_rm, id), proc_data_imputed)
```






