---
title: "Project title"
author:
  - name: Author Name
    affiliation:
    - &aff1 Author Institution, City, Country.
    email: email@website.com
date: "2026-02-23"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
vignette: >
  %\VignetteIndexEntry{Project title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Programming environment

```{r Set random seed, include=FALSE}
seed <- 2026-02-23
```

```{r Load packages, include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(ggpubr)
library(pbapply)
library(broom)
library(brms)
library(cmdstanr) # Comment this if stand is not installable
extract <- tidyr::extract
library(broom.mixed) # Comment this if stand is not installable
library(mice)
filter <- dplyr::filter
cbind <- base::cbind
rbind <- base::rbind
library(igraph)
library(ggnetwork)
library(cluster)
```

```{r Load functions, include=FALSE}
lapply(list.files("R/", full.names = TRUE), source)
```

```{r Set theme, include=FALSE}
my_theme_set()
```

```{r Session information, echo=FALSE}
sessionInfo()
```

# Data Preprocessing

## Raw data

```{r Read code book, include=FALSE}
code_book <- read_csv("inst/extdata/code_book.csv", show_col_types = FALSE)
```

```{r Show code book, echo=FALSE}
show_table(code_book, "code_book", "Code book.")
```

```{r Column-type string, include=FALSE}
coltype_str <- paste0(code_book$col_type, collapse = "")
```

```{r Raw data, include=FALSE}
raw_data <- read_csv("inst/extdata/raw_data.csv", col_types = coltype_str)
```

```{r State row and column numbers of raw data, echo=FALSE}
cat("Rows: ", nrow(raw_data), "\n")
cat("Columns: ", ncol(raw_data))
```

## Select variable of interest

```{r Variable of interest, include=FALSE}
voi <- "pe"
```

```{r Rename the variable as outcome and put after identifier, include=FALSE}
proc_data <-
  raw_data |>
  rename(outcome = voi) |>
  select(id, outcome, everything())
```

# Data partition

```{r Determine partition sample size, include=FALSE}
whole_id <- proc_data$id
whole_n <- length(whole_id)
test_n <- ceiling(0.2 * whole_n)
dev_n <- whole_n - test_n
val_n <- ceiling(0.2 * dev_n)
train_n <- dev_n - val_n
```

```{r Randomly select identifiers for each partition, include=FALSE}
set.seed(seed)
test_id <- sample(whole_id, test_n)
dev_id <- setdiff(whole_id, test_id)

set.seed(seed)
val_id <- sample(dev_id, val_n)
train_id <- setdiff(dev_id, val_id)
```

```{r Create set data, include=FALSE}
set_data <-
  data.frame(id = train_id, set = "training") |>
  rbind(data.frame(id = val_id, set = "validation")) |>
  rbind(data.frame(id = test_id, set = "testing")) |>
  mutate(set = factor(set, c("training", "validation", "testing")))
```

```{r Join set data to outcome data, include=FALSE}
outcome_set_data <-
  proc_data |>
  select(id, outcome) |>
  left_join(set_data, by = "id")
```

```{r Summarize outcome class per set, include=FALSE}
if(class(outcome_set_data$outcome) == "numeric"){
  outcome_set_summary <-
    outcome_set_data |>
    group_by(set) |>
    summarize(n = n(), .groups = "drop")
  
  outcome_set_plot <-
    outcome_set_summary |>
    ggplot(aes(set, n)) +
    geom_col()
}else{
  outcome_set_summary <-
    outcome_set_data |>
    group_by(set, outcome) |>
    summarize(n = n(), .groups = "drop")
  
  outcome_set_plot <-
    outcome_set_summary |>
    ggplot(aes(set, n, fill = outcome)) +
    geom_col()
}
```

```{r Show summary per set, echo=FALSE, fig.height=90/25.4, fig.width=190/25.4}
outcome_set_plot
```

```{r Caption outcome class summary per set, echo=FALSE}
show_figure_legend("outcome_set_plot", "Data partition.")
```

```{r Split samples using the identifiers for train set, include=FALSE}
train_data <- slice(proc_data, match(train_id, id))
```

# QQ plot

```{r Select numerical variables, include=FALSE}
num_data <- select_if(train_data, is.numeric)
```

```{r State number of QQ plots, echo=FALSE}
cat("Number of QQ plots (= # of numeric variables): ", ncol(num_data))
```

```{r Create custom function for QQ plot}
qq_plotter <- function(variable, data){
  data |>
    `colnames<-`(str_replace_all(colnames(data), variable, "variable")) |>
    ggplot(aes(sample = variable)) +
  	geom_qq() +
  	geom_qq_line() +
    xlab("Normal +/- SD") +
    ylab(variable) +
    theme(axis.title = element_text(size = 10))
}
```

```{r Create QQ plots, include=FALSE}
if(ncol(num_data) > 0){
  qq_plots <- pblapply(colnames(num_data), qq_plotter, num_data)
  
  qq_ncol <- 4
  qq_nrow <- ceiling(length(qq_plots) / qq_ncol)
  qq_plots <- ggarrange(plotlist = qq_plots, ncol = qq_ncol, nrow = qq_nrow)
}
```

```{r Show QQ plots, echo=FALSE, fig.height=270/25.4, fig.width=190/25.4}
if(ncol(num_data) > 0){
  qq_plots
}
```

```{r Caption QQ plots, echo=FALSE}
if(ncol(num_data) > 0){
  show_figure_legend("qq_plots", "QQ plots.")
}
```

# Data transformation

```{r Variables for squared-root test-transformation, include=FALSE}
ttrans_var_sqrt <- c("plgf", "sflt1_plgf")
```

```{r Variables for squared test-transformation, include=FALSE}
ttrans_var_sq <- c()
```

```{r Variables for log test-transformation, include=FALSE}
ttrans_var_log <- c()
```

```{r State test-ransformed variables, echo=FALSE}
ttrans_var <- c(ttrans_var_sqrt, ttrans_var_sq, ttrans_var_log)

cat(
  paste0("Test-transformed variables (", length(ttrans_var), "): ")
  , ttrans_var
)
```

```{r Create custom function for squared transformation}
sq <- function(x){
  x^2
}
```

```{r Data test-transformation using numerical train set}
if(ncol(num_data) > 0 & length(ttrans_var) > 0){
  num_data_ttrans <-
    num_data |>
    mutate_at(ttrans_var_sqrt, sqrt) |>
    mutate_at(ttrans_var_sq, sq) |>
    mutate_at(ttrans_var_log, log)
}
```

```{r Create QQ test-trans, include=FALSE}
if(ncol(num_data_ttrans) > 0 & length(ttrans_var) > 0){
  qt_plots <- pblapply(ttrans_var, qq_plotter, num_data_ttrans)
  
  qt_ncol <- 4
  qt_nrow <- ceiling(length(qt_plots) / qt_ncol)
  qt_plots <- ggarrange(plotlist = qt_plots, ncol = qt_ncol, nrow = qt_nrow)
}
```

```{r Show QQ test-trans, echo=FALSE, fig.height=54/25.4, fig.width=190/25.4}
if(ncol(num_data_ttrans) > 0 & length(ttrans_var) > 0){
  qt_plots
}
```

```{r Caption QQ test-trans, echo=FALSE}
if(ncol(num_data_ttrans) > 0 & length(ttrans_var) > 0){
  show_figure_legend("qt_plots", "QQ plots after data transformation.")
}
```

```{r Variables for squared-root transformation, include=FALSE}
trans_var_sqrt <- c("sflt1_plgf")
```

```{r Variables for squared transformation, include=FALSE}
trans_var_sq <- c()
```

```{r Variables for log transformation, include=FALSE}
trans_var_log <- c()
```

```{r State transformed variables, echo=FALSE}
trans_var <- c(trans_var_sqrt, trans_var_sq, trans_var_log)
cat(paste0("Transformed variables (", length(trans_var), "): "), trans_var)
```

```{r State categorized numerical variables, echo=FALSE}
cat_num_var <- setdiff(ttrans_var, trans_var)
cat_num_var <- c(cat_num_var, "uta_bn", "parity")
cat(
  paste0("Categorized numerical variables (", length(cat_num_var), "): ")
  , cat_num_var
)
```

```{r Create custom function categorize column using pre-defined cutoff}
categorize_from_table <- function(data, cutoff_table){
  
  if(nrow(cutoff_table) == 0) return(data)
  
  cols <- unique(cutoff_table$column)
  
  for(col_name in cols){
    
    rules <-
      cutoff_table |>
      filter(column == col_name) |>
      arrange(cutoff)
    
    breaks <- c(rules$cutoff, Inf)
    labels <- rules$category
    
    data[[col_name]] <-
      cut(
        data[[col_name]]
        , breaks = breaks
        , labels = labels
        , right = FALSE
        , include.lowest = TRUE
      )
  }
  
  mutate_at(data, cols, factor)
}
```

```{r Read pre-defined cutoff for categorization, include=FALSE}
category_cutoff <-
  read_csv("inst/extdata/category_cutoff.csv", show_col_types = FALSE)
```

```{r Transform data for all sets, include=FALSE}
if(ncol(num_data) > 0){
  
  proc_data_trans <- proc_data
  
  # Numerical data transformation
  if(length(trans_var) > 0){
    proc_data_trans <-
      proc_data_trans |>
      mutate_at(trans_var_sqrt, sqrt) |>
      mutate_at(trans_var_sq, sq) |>
      mutate_at(trans_var_log, log)
  }
  
  # Numerical data categorization
  code_book_trans <- code_book
  if(length(intersect(cat_num_var, unique(category_cutoff$column))) > 0){
    proc_data_trans <-
      proc_data_trans |>
      categorize_from_table(category_cutoff)
    
    code_book_trans <-
      read_csv("inst/extdata/code_book_trans.csv", show_col_types = FALSE)
  }
}
```

```{r Obtain train set from transformed dataset, include=FALSE}
train_data_trans <- slice(proc_data_trans, match(train_id, id))
```

# Determine Tukey's fences

```{r Select numerical variables from transformed dataset, include=FALSE}
num_data_trans <- select_if(train_data_trans, is.numeric)
```

```{r State number of numeric variables after transformation, echo=FALSE}
cat("Number of numeric variables after transformation: ", ncol(num_data_trans))
```

```{r Create custom function to determine Tukey fences}
tukey_fence <- function(x, which = c("lb", "ub"), na.rm = TRUE){
	which <- match.arg(which)

	q1 <- quantile(x, 0.25, na.rm = na.rm)
	q3 <- quantile(x, 0.75, na.rm = na.rm)
	iqr <- q3 - q1

	if(which == "lb"){
		q1 - 1.5 * iqr
	}else{
		q3 + 1.5 * iqr
	}
}
```

```{r Determine Tukey fences from training set, include=FALSE}
if(ncol(num_data_trans) > 0){
  tukey_fence_lb <-
    num_data_trans |>
    summarize_all(tukey_fence, which = "lb") |>
    gather(col_name, tukey_lb)
  
  tukey_fence_ub <-
    num_data_trans |>
    summarize_all(tukey_fence, which = "ub") |>
    gather(col_name, tukey_ub)
  
  tukey_fences <-
    tukey_fence_lb |>
    left_join(tukey_fence_ub, by = join_by(col_name))
}
```

# Outliers handling 

```{r Flag outliers using Tukey fences for all sets, include=FALSE}
outliers <-
  proc_data_trans |>
  select(colnames(num_data_trans)) |>
  mutate(seq = seq(n())) |>
  gather(col_name, value, -seq) |>
  left_join(tukey_fences, by = join_by(col_name)) |>
  mutate(is_outlier = value < tukey_lb | value > tukey_ub) |>
  select(seq, col_name, is_outlier) |>
  spread(col_name, is_outlier) |>
  arrange(seq) |>
  select(colnames(num_data_trans))
```

```{r Summarize outliers for train set, include=FALSE}
outliers_sum <-
  select(proc_data_trans, id) |>
  cbind(outliers) |>
  slice(match(train_id, id)) |>
  select(-id) |>
  summarize_all(mean, na.rm = TRUE) |>
  gather(variable, p) |>
  mutate(p = round(p * 100, 2))
```

```{r Show outliers summary, echo=FALSE}
show_table(outliers_sum, "outliers_sum", "Outliers summary in train set.")
```

```{r Remove outliers by assigning values to missing, include=FALSE}
num_proc_data_outlier_rm <-
  proc_data_trans |>
  select(colnames(num_data_trans))

num_proc_data_outlier_rm[as.matrix(outliers)] <- NA
```

```{r Bind back outlier-removed data for all sets, include=FALSE}
proc_data_outlier_rm <-
  proc_data_trans |>
  select(setdiff(colnames(proc_data_trans), colnames(num_data_trans))) |>
  cbind(num_proc_data_outlier_rm) |>
  select(colnames(proc_data_trans))
```

```{r Obtain train set from outlier-removed dataset, include=FALSE}
train_data_outlier_rm <- slice(proc_data_outlier_rm, match(train_id, id))
```

```{r Add missingness indicators, include=FALSE}
train_data_ms <-
  train_data_outlier_rm |>
  select_if(\(x) any(is.na(x))) |>
  mutate_all(is.na) |>
  mutate_all(as.numeric) |>
  mutate_all(factor) |>
  rename_all(paste0, "_ms")

train_data_ms <-
  train_data_outlier_rm |>
  cbind(train_data_ms)
```

# Normality assessment and test

```{r Assess normality by QQ plot, include=FALSE}
normal_qq <- c("age", "height")
```

```{r State normal numerical variables by QQ plots, echo=FALSE}
cat("Normal numerical variables by QQ plot: ",normal_qq)
```

```{r List non-normal variables by QQ plot, include=FALSE}
non_normal_qq <- setdiff(colnames(num_data_trans), normal_qq)
```

```{r Conduct Shapiro-Wilk normality test, include=FALSE}
sw_results <-
  num_data_trans |>
  select_at(non_normal_qq) |>
  lapply(shapiro.test) |>
  lapply(tidy) |>
  imap(~ cbind(variable = .y, .x)) |>
  reduce(rbind) |>
  select(-method)
```

```{r Show Shapiro-Wilk normality test, echo=FALSE}
show_table(sw_results, "sw_results", "Shapiro-Wilk normality test results.")
```

```{r List normal variables by Shapiro-Wilk normality test, include=FALSE}
normal_sw <-
  sw_results |>
  filter(p.value > 0.05) |>
  pull(variable)
```

```{r List non-normal variables by Shapiro-Wilk normality test, include=FALSE}
non_normal_sw <- setdiff(non_normal_qq, normal_sw)
```

```{r Obtain normality for numerical columns, include=FALSE}
normality_data <-
  data.frame(col_name = colnames(num_data_trans)) |>
  mutate(normal = col_name %in% c(normal_qq, normal_sw))
```

# Perfect/partial separation

```{r Obtain scale for each column name except identifier, include=FALSE}
scale_data <-
  data.frame(
     scale =
       train_data_ms |>
       select(-id) |>
       sapply(class)
   ) |>
  rownames_to_column(var = "col_name")
```

```{r Identify all possible pair combinations, include=FALSE}
comb_pair <-
  combn(scale_data$col_name, 2) |>
  t() |>
  as.data.frame() |>
  left_join(
    rename(scale_data, V1 = col_name, scale_V1 = scale)
    , by = join_by(V1)
  ) |>
  left_join(
    rename(scale_data, V2 = col_name, scale_V2 = scale)
    , by = join_by(V2)
  )
```

```{r Create custom function to summarize variable pairs}
pair_summary <- function(data, V1, V2){
  result <-
    data |>
    group_by_at(c(V1, V2)) |>
    summarize(n = n(), .groups = "drop") |>
    `colnames<-`(c("value_V1", "value_V2", "n"))
  
  expand.grid(value_V1 = levels(data[[V1]]), value_V2 = levels(data[[V2]])) |>
    left_join(result, by = join_by(value_V1, value_V2)) |>
    mutate(n = ifelse(is.na(n), 0, n)) |>
    mutate(V1 = !!V1, V2 = !!V2)
}
```

```{r Perfect or partial separation data for categorical pairs, include=FALSE}
ps_data <-
  comb_pair |>
  filter(scale_V1 == "factor" & scale_V2 == "factor") |>
  select(V1, V2) |>
  pmap(\(V1, V2) pair_summary(train_data_ms, V1, V2)) |>
  reduce(rbind) |>
  group_by(V1, V2) |>
  summarize(ps = sum(n == 0) > 0, .groups = "drop")

# Temporarily add this if stan is not installable
# |>
#   mutate(ps = FALSE) 
```

# Correlation matrix

```{r Obtain number of categories for categorical variables, include=FALSE}
level_data <-
  train_data_ms |>
  select_if(is.factor) |>
  summarize_all(\(x) length(intersect(levels(x), unique(x)))) |>
  gather(col_name, level_n)
```

```{r Identify untestable pairs, include=FALSE}
comb_pair_untestable <-
  comb_pair |>
  filter(str_remove_all(V1, "_ms$") == str_remove_all(V2, "_ms$"))
```

```{r Join combinations with normality-separation-level data, include=FALSE}
comb_pair_normality_ps_level <-
  comb_pair |>
  anti_join(comb_pair_untestable, by = join_by(V1, V2, scale_V1, scale_V2)) |>
  left_join(
    rename(normality_data, V1 = col_name, normal_V1 = normal)
    , by = join_by(V1)
  ) |>
  left_join(
    rename(normality_data, V2 = col_name, normal_V2 = normal)
    , by = join_by(V2)
  ) |>
  left_join(ps_data, by = join_by(V1, V2))  |>
  left_join(
    rename(level_data, V1 = col_name, level_V1 = level_n)
    , by = join_by(V1)
  ) |>
  left_join(
    rename(level_data, V2 = col_name, level_V2 = level_n)
    , by = join_by(V2)
  ) |>
  mutate(
    normal_V1 = ifelse(is.na(normal_V1), TRUE, normal_V1)
    , normal_V2 = ifelse(is.na(normal_V2), TRUE, normal_V2)
    , ps = ifelse(is.na(ps), FALSE, ps)
    , level_V1 = ifelse(is.na(level_V1), 0, level_V1)
    , level_V2 = ifelse(is.na(level_V2), 0, level_V2)
  )
```

## Numerical-to-numerical pairs

```{r Identify numerical-to-numerical pairs, include=FALSE}
comb_pair_num_num <-
  comb_pair_normality_ps_level |>
  filter(scale_V1 == "numeric" & scale_V2 == "numeric")
```

```{r Identify normal-to-normal pairs, include=FALSE}
comb_pair_normal_normal <-
  comb_pair_num_num |>
  filter(normal_V1 & normal_V2) |>
  select(V1, V2)
```

```{r Identify non-normal-to-normal pairs, include=FALSE}
comb_pair_non_normal_normal <-
  comb_pair_num_num |>
  select(V1, V2) |>
  anti_join(comb_pair_normal_normal, join_by(V1, V2))
```

```{r Create custom function for correlation test}
correlation_test <- function(data, V1, V2, method){
  cor.test(train_data_ms[[V1]], train_data_ms[[V2]], method = "pearson") |>
    tidy() |>
    mutate(V1 = !!V1, V2 = !!V2) |>
    select(V1, V2, everything())
}
```

```{r Conduct Pearson correlation tests, include=FALSE}
if(nrow(comb_pair_normal_normal) > 0){
  cm_pearson <-
    comb_pair_normal_normal |>
    pmap(\(V1, V2) correlation_test(train_data_ms, V1, V2, "pearson")) |>
    reduce(rbind)
}
```

```{r Conduct Spearman correlation tests, include=FALSE}
if(nrow(comb_pair_non_normal_normal) > 0){
  cm_spearman <-
    comb_pair_non_normal_normal |>
    pmap(\(V1, V2) correlation_test(train_data_ms, V1, V2, "spearman")) |>
    reduce(rbind)
}
```

## Categorical-to-categorical pairs

```{r Identify categorical-to-categorical pairs, include=FALSE}
comb_pair_cat_cat <-
  comb_pair_normality_ps_level |>
  filter(scale_V1 == "factor" & scale_V2 == "factor")
```

```{r Identify pairs with separation, include=FALSE}
comb_pair_ps <-
  comb_pair_cat_cat |>
  filter(ps)
```

```{r Identify separation pairs with >2-category V1, include=FALSE}
comb_pair_ps_gt2_cat_V1 <-
  comb_pair_ps |>
  filter(level_V1 > 2) |>
  select(V1, V2)
```

```{r Create custom function for Bayesian categorical regression}
bayesian_cat_reg_fast <- function(data, V1, V2){
  data2 <-
    data |>
    select(all_of(c(V1, V2))) |>
    drop_na() |>
    droplevels()

  form <- reformulate(termlabels = V2, response = V1)

  result<-
    brm(
      formula = form
      , family = categorical()
      , data = data2
      , backend = "cmdstanr"
      , chains = 1
      , iter = 500
      , warmup = 250
      , refresh = 0
      , silent = 2
      , algorithm = "pathfinder"
    )
  
  tidy(result) |>
    mutate(V1 = !!V1, V2 = !!V2) |>
    select(V1, V2, everything())
}
```

```{r Conduct Bayesian categorical regression, include=FALSE}
if(nrow(comb_pair_ps_gt2_cat_V1) > 0){
  cm_bayesian_cat_reg <-
    comb_pair_ps_gt2_cat_V1 |>
    pmap(\(V1, V2) bayesian_cat_reg_fast(train_data_ms, V1, V2)) |>
    reduce(rbind)
}
```

```{r Identify separation pairs with 2-category V1, include=FALSE}
comb_pair_ps_2_cat_V1 <-
  comb_pair_ps |>
  filter(level_V1 == 2) |>
  select(V1, V2)
```

```{r Create custom function for Bayesian Bernoulli logistic regression}
bayesian_bern_reg_fast <- function(data, V1, V2){
  data2 <-
    data |>
    select(all_of(c(V1, V2))) |>
    drop_na() |>
    droplevels()

  form <- reformulate(termlabels = V2, response = V1)

  result <-
    brm(
      formula = form
      , family = bernoulli()
      , data = data2
      , backend = "cmdstanr"
      , chains = 1
      , iter = 500
      , warmup = 250
      , refresh = 0
      , silent = 2
      , algorithm = "pathfinder"
    )
  
  tidy(result) |>
    mutate(V1 = !!V1, V2 = !!V2) |>
    select(V1, V2, everything())
}
```

```{r Conduct Bayesian Bernoulli logistic regression, include=FALSE}
if(nrow(comb_pair_ps_2_cat_V1) > 0){
  cm_bayesian_bern_reg <-
    comb_pair_ps_2_cat_V1 |>
    pmap(\(V1, V2) bayesian_bern_reg_fast(train_data_ms, V1, V2)) |>
    reduce(rbind)
}
```

```{r Identify pairs without separation, include=FALSE}
comb_pair_non_ps <-
  comb_pair_cat_cat |>
  filter(!ps) |>
  select(V1, V2)
```

```{r Create custom function for Fisher exact test}
fisher_exact_test_fast <- function(data, V1, V2){
  result <-
    fisher.test(
      data[[V1]], data[[V2]]
      , workspace = 1e7, simulate.p.value = TRUE
    )
  
  tidy(result) |>
    select(p.value, method, alternative) |>
    mutate(V1 = !!V1, V2 = !!V2) |>
    select(V1, V2, everything())
}
```

```{r Conduct Fisher exact test, include=FALSE}
if(nrow(comb_pair_non_ps) > 0){
  cm_fisher <-
    comb_pair_non_ps |>
    pmap(\(V1, V2) fisher_exact_test_fast(train_data_ms, V1, V2)) |>
    reduce(rbind)
}
```

## Mixed-scale pairs

```{r Identify categorical-to-numerical pairs, include=FALSE}
comb_pair_cat_num <-
  comb_pair_normality_ps_level |>
  filter(scale_V1 == "factor" & scale_V2 == "numeric")
```

```{r Identify >2-categorical-to-norm-numerical pairs, include=FALSE}
comb_pair_cat_gt2_num_norm <-
  comb_pair_cat_num |>
  filter(level_V1 > 2) |>
  filter(normal_V2) |>
  select(V1, V2) |>
  mutate(inv = FALSE)
```

```{r Identify >2-categorical-to-non-norm-numerical pairs, include=FALSE}
comb_pair_cat_gt2_num_non_norm <-
  comb_pair_cat_num |>
  filter(level_V1 > 2) |>
  filter(!normal_V2) |>
  select(V1, V2) |>
  mutate(inv = FALSE)
```

```{r Identify 2-categorical-to-norm-numerical pairs, include=FALSE}
comb_pair_cat_2_num_norm <-
  comb_pair_cat_num |>
  filter(level_V1 == 2) |>
  filter(normal_V2) |>
  select(V1, V2) |>
  mutate(inv = FALSE)
```

```{r Identify 2-categorical-to-non-norm-numerical pairs, include=FALSE}
comb_pair_cat_2_num_non_norm <-
  comb_pair_cat_num |>
  filter(level_V1 == 2) |>
  filter(!normal_V2) |>
  select(V1, V2) |>
  mutate(inv = FALSE)
```

```{r Identify numerical-to-categorical pairs, include=FALSE}
comb_pair_num_cat <-
  comb_pair_normality_ps_level |>
  filter(scale_V1 == "numeric" & scale_V2 == "factor")
```

```{r Identify norm-numerical-to->2-categorical pairs, include=FALSE}
comb_pair_num_norm_cat_gt2 <-
  comb_pair_num_cat |>
  filter(level_V2 > 2)  |>
  filter(normal_V1) |>
  select(V1 = V2, V2 = V1) |>
  mutate(inv = TRUE)
```

```{r Identify non-norm-numerical-to->2-categorical pairs, include=FALSE}
comb_pair_num_non_norm_cat_gt2 <-
  comb_pair_num_cat |>
  filter(level_V2 > 2)  |>
  filter(!normal_V1) |>
  select(V1 = V2, V2 = V1) |>
  mutate(inv = TRUE)
```

```{r Identify norm-numerical-to-2-categorical pairs, include=FALSE}
comb_pair_num_norm_cat_2 <-
  comb_pair_num_cat |>
  filter(level_V2 == 2)  |>
  filter(normal_V1) |>
  select(V1 = V2, V2 = V1) |>
  mutate(inv = TRUE)
```

```{r Identify non-norm-numerical-to-2-categorical pairs, include=FALSE}
comb_pair_num_non_norm_cat_2 <-
  comb_pair_num_cat |>
  filter(level_V2 == 2)  |>
  filter(!normal_V1) |>
  select(V1 = V2, V2 = V1) |>
  mutate(inv = TRUE)
```

```{r Join >2-categorical and norm-numerical pairs, include=FALSE}
comb_pair_mixed_gt2_norm <-
  rbind(comb_pair_cat_gt2_num_norm, comb_pair_num_norm_cat_gt2)
```

```{r Create custom function for one-way ANOVA test}
one_way_anova_test <- function(data, V1, V2, inv){
  data <- data.frame(V1 = data[[V1]], V2 = data[[V2]])
  
  result <- aov(V2 ~ V1, data = data)
  
  result <-
    tidy(result) |>
    mutate(V1 = !!V1, V2 = !!V2)
  
  if(inv){
    result |>
      mutate(term = ifelse(term == "V1", "V2", term)) |>
      select(V1 = V2, V2 = V1, everything())
  }else{
    result |>
      select(V1, V2, everything())
  }
}
```

```{r Conduct one-way ANOVA test, include=FALSE}
if(nrow(comb_pair_mixed_gt2_norm) > 0){
  cm_one_way_anova <-
    comb_pair_mixed_gt2_norm |>
    pmap(\(V1, V2, inv) one_way_anova_test(train_data_ms, V1, V2, inv)) |>
    reduce(rbind)
}
```

```{r Join >2-categorical and non-norm-numerical pairs, include=FALSE}
comb_pair_mixed_gt2_non_norm <-
  rbind(comb_pair_cat_gt2_num_non_norm, comb_pair_num_non_norm_cat_gt2)
```

```{r Create custom function for Kruskal-Wallis test}
kruskal_wallis_test <- function(data, V1, V2, inv){
  V2_group_V1 <-
    data[[V1]] |>
    unique() |>
    lapply(\(x) data[[V2]][data[[V1]] == x])
  
  result <- kruskal.test(V2_group_V1)
  
  result <-
    tidy(result) |>
    mutate(V1 = !!V1, V2 = !!V2)
  
  if(inv){
    result |>
      select(V1 = V2, V2 = V1, everything())
  }else{
    result |>
      select(V1, V2, everything())
  }
}
```

```{r Conduct Kruskal-Wallis test, include=FALSE}
if(nrow(comb_pair_mixed_gt2_non_norm) > 0){
  cm_kruskal_wallis <-
    comb_pair_mixed_gt2_non_norm |>
    pmap(\(V1, V2, inv) kruskal_wallis_test(train_data_ms, V1, V2, inv)) |>
    reduce(rbind)
}
```

```{r Join 2-categorical and norm-numerical pairs, include=FALSE}
comb_pair_mixed_2_norm <-
  rbind(comb_pair_cat_2_num_norm, comb_pair_num_norm_cat_2)
```

```{r Create custom function for Welch two-sample t-test}
welch_two_sample_t_test <- function(data, V1, V2, inv){
  V2_group_V1_1 = data[[V2]][data[[V1]] == unique(data[[V1]])[1]]
  V2_group_V1_2 = data[[V2]][data[[V1]] == unique(data[[V1]])[2]]
  
  result <-
    try(
      t.test(
        V2_group_V1_1
        , V2_group_V1_2
        , alternative = "two.sided"
        , paired = FALSE
        , var.equal = FALSE
      )
    )
  
  if(is.character(result)){
    result <-
      data.frame(
        estimate = as.numeric(NA)
        , estimate1 = as.numeric(NA)
        , estimate2 = as.numeric(NA)
        , statistic = as.numeric(NA)
        , p.value = as.numeric(NA)
        , parameter = as.numeric(NA)
        , conf.low = as.numeric(NA)
        , conf.high = as.numeric(NA)
        , method =
          paste0(
            "Welch Two Sample t-test ("
            , paste0(result, collapse = ". ")
            , ")"
          )
        , alternative = "two.sided"
      )
  }else{
    result <- tidy(result)
  }
  
  result <- mutate(result, V1 = !!V1, V2 = !!V2)
  
  if(inv){
    result |>
      select(V1 = V2, V2 = V1, everything())
  }else{
    result |>
      select(V1, V2, everything())
  }
}
```

```{r Conduct Welch two-sample t-test, include=FALSE}
if(nrow(comb_pair_mixed_2_norm) > 0){
  cm_welch_two_sample_t <-
    comb_pair_mixed_2_norm |>
    pmap(\(V1, V2, inv) welch_two_sample_t_test(train_data_ms, V1, V2, inv)) |>
    reduce(rbind)
}
```

```{r Join 2-categorical and non-norm-numerical pairs, include=FALSE}
comb_pair_mixed_2_non_norm <-
  rbind(comb_pair_cat_2_num_non_norm, comb_pair_num_non_norm_cat_2)
```

```{r Create custom function for Wilcoxon rank-sum test}
wilcoxon_rank_sum_test <- function(data, V1, V2, inv){
  V2_group_V1_1 = data[[V2]][data[[V1]] == unique(data[[V1]])[1]]
  V2_group_V1_2 = data[[V2]][data[[V1]] == unique(data[[V1]])[2]]
  
  result <-
    try(
      wilcox.test(
        V2_group_V1_1
        , V2_group_V1_2
        , alternative = "two.sided"
        , paired = FALSE
      )
    )
  
  if(is.character(result)){
    result <-
      data.frame(
        statistic = as.numeric(NA)
        , p.value = as.numeric(NA)
        , method =
          paste0(
            "Welch Two Sample t-test ("
            , paste0(result, collapse = ". ")
            , ")"
          )
        , alternative = "two.sided"
      )
  }else{
    result <- tidy(result)
  }
  
  result <- mutate(result, V1 = !!V1, V2 = !!V2)
  
  if(inv){
    result |>
      select(V1 = V2, V2 = V1, everything())
  }else{
    result |>
      select(V1, V2, everything())
  }
}
```

```{r Conduct Welch two-sample t-test, include=FALSE}
if(nrow(comb_pair_mixed_2_non_norm) > 0){
  cm_wilcoxon_rank_sum <-
    comb_pair_mixed_2_non_norm |>
    pmap(\(V1, V2, inv) wilcoxon_rank_sum_test(train_data_ms, V1, V2, inv)) |>
    reduce(rbind)
}
```

## Multiple testing correction

```{r Create custom function to translate CI to p-value}
translate_ci_to_p_value <- function(data, V1, V2){
  data |>
    filter(V1 == !!V1 & V2 == !!V2) |>
    filter(!str_detect(term, "\\(Intercept\\)")) |>
    filter(!(conf.low <= 0 & conf.high >= 0)) |>
    summarize(n_sig = n()) |>
    mutate(p.value = ifelse(n_sig > 0, 0, 1)) |>
    filter(!is.na(p.value)) |>
    mutate(V1 = V1, V2 = V2) |>
    select(V1, V2, p.value)
}
```

```{r Translate CI to p-value for Bayesian categorical regression, include=FALSE}
cm_bayesian_cat_reg_p_value <-
  cm_bayesian_cat_reg |>
  select(V1, V2) |>
  unique() |>
  pmap(\(V1, V2) translate_ci_to_p_value(cm_bayesian_cat_reg, V1, V2)) |>
  reduce(rbind) |>
  mutate(method = "Bayesian categorical regression")
```

```{r Translate CI to p-value for Bayesian Bernoulli log reg, include=FALSE}
cm_bayesian_bern_reg_p_value <-
  cm_bayesian_bern_reg |>
  select(V1, V2) |>
  unique() |>
  pmap(\(V1, V2) translate_ci_to_p_value(cm_bayesian_bern_reg, V1, V2)) |>
  reduce(rbind) |>
  mutate(method = "Bayesian Bernoulli logistic regression")
```

```{r Combine results from all correlation tests, include=FALSE}
correlation_results <- list()

if(nrow(comb_pair_untestable) > 0){
  correlation_results <-
    correlation_results |>
    c(list(
        untestable =
          comb_pair_untestable |>
          select(V1, V2) |>
          mutate(p.value = as.numeric(NA), method = "None (untestable)")
      )
    )
}

if(nrow(comb_pair_normal_normal) > 0){
  correlation_results <- c(correlation_results, list(pearson = cm_pearson))
}

if(nrow(comb_pair_non_normal_normal) > 0){
  correlation_results <- c(correlation_results, list(spearman = cm_spearman))
}

if(nrow(comb_pair_ps_gt2_cat_V1) > 0){
  correlation_results <-
    c(correlation_results, list(bayesian_cat_reg = cm_bayesian_cat_reg_p_value))
}

if(nrow(comb_pair_ps_2_cat_V1) > 0){
  correlation_results <-
    correlation_results |>
    c(list(bayesian_bern_reg = cm_bayesian_bern_reg_p_value))
}

if(nrow(comb_pair_non_ps) > 0){
  correlation_results <- c(correlation_results, list(fisher = cm_fisher))
}

if(nrow(comb_pair_mixed_gt2_norm) > 0){
  correlation_results <-
    correlation_results |>
    c(list(
        one_way_anova =
          cm_one_way_anova |>
          filter(term != "Residuals") |>
          mutate(method = "One-way ANOVA")
      )
    )
}

if(nrow(comb_pair_mixed_gt2_non_norm) > 0){
  correlation_results <-
    c(correlation_results, list(kruskal_wallis = cm_kruskal_wallis))
}

if(nrow(comb_pair_mixed_2_norm) > 0){
  correlation_results <-
    c(correlation_results, list(welch_two_sample_t = cm_welch_two_sample_t))
}

if(nrow(comb_pair_mixed_2_non_norm) > 0){
  correlation_results <-
    c(correlation_results, list(wilcoxon_rank_sum = cm_wilcoxon_rank_sum))
}
```

```{r Combine p-values from all correlation tests, include=FALSE}
correlation_p_values <-
  correlation_results |>
  lapply(select, V1, V2, p.value, method) |>
  reduce(rbind)
```

```{r Compute FDRs from p-values by Benjamini-Hochberg method, include=FALSE}
correlation_fdr <-
  correlation_p_values |>
  mutate(
    FDR =
      p.value |>
      p.adjust(method = "BH")
    , sig =
      ifelse(is.na(FDR), "Not tested", ifelse(FDR <= 0.05, "Yes", "No")) |>
      factor()
    , method_no =
      as.numeric(factor(method, unique(method))) |>
      factor()
  ) |>
  select(sig, FDR, everything())
```

## Visualize correlation by significance

```{r Create custom function to plot correlation matrix}
cor_matrix_plotter <- function(data, triangle = "lower"){
  vars <- union(as.character(data$V1), as.character(data$V2))

  data_plot <-
    data |>
    mutate(
      V1 = factor(V1, levels = vars),
      V2 = factor(V2, levels = rev(vars)),
      sig = factor(sig, levels = c("No", "Yes", "Not tested"))
    ) |>
    mutate(
      x_id = as.integer(factor(V1, levels = vars)),
      y_var = as.character(V2),
      y_id = match(y_var, rev(vars))
    )

  if(triangle == "lower"){
    data_plot <- data_plot |> filter(x_id + y_id <= length(vars) + 1)
  }else if(triangle == "upper"){
    data_plot <- data_plot |> filter(x_id + y_id >= length(vars) + 1)
  }

  ggplot(data_plot, aes(x = V1, y = V2, fill = sig)) +
    geom_tile(color = "white") +
    geom_text(aes(label = method_no), size = 3) +
    scale_fill_manual(
      values = c(
        "No" = "grey80",
        "Yes" = "steelblue",
        "Not tested" = "grey40"
      )
    ) +
    coord_equal() +
    theme_minimal() +
    theme(
      panel.grid = element_blank(),
      axis.title = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
    labs(fill = "Significant")
}
```

```{r Plot cor matrix, include=FALSE}
correlation_matrix <-
  comb_pair |>
  select(V1, V2) |>
  left_join(correlation_fdr, by = join_by(V1, V2)) |>
  cor_matrix_plotter()
```

```{r Show corr matrix, echo=FALSE, fig.height=150/25.4, fig.width=190/25.4}
correlation_matrix
```

```{r Caption corr matrix, echo=FALSE}
show_figure_legend("correlation_matrix", "Correlation matrix.")
```

```{r Obtain correlation method numbers, include=FALSE}
cm_method_no <-
  correlation_matrix$data |>
  select(no = method_no, method) |>
  unique() |>
  arrange(no)
```

```{r Show correlation method numbers, echo=FALSE}
show_table(cm_method_no, "cm_method_no", "Correlation methods.")
```

# Missing value imputation

```{r Create custom function to make MICE predictor matrix, include=FALSE}
make_mice_pred_matrix <- function(
  corr_data
  , data_names = NULL
  , sig_value = "Yes"
  , ms_suffix = "_ms"
){
  # Variables only, exclude missingness indicators
  if(is.null(data_names)){
    data_names <-
      union(
        as.character(corr_data$V1)
        , as.character(corr_data$V2)
      )

    data_names <-
      data_names[!grepl(paste0(ms_suffix, "$"), data_names)] |>
      sort()
  }

  pred_matrix <-
    matrix(
      0
      , nrow = length(data_names)
      , ncol = length(data_names)
      , dimnames = list(data_names, data_names)
    )

  corr_yes <-
    corr_data |>
    filter(sig == sig_value) |>
    select(V1, V2) |>
    distinct()

  for(v in data_names){
    ms_v <- paste0(v, ms_suffix)

    preds_from_v <-
      corr_yes |>
      filter(V1 == v) |>
      pull(V2) |>
      as.character()

    preds_from_ms <-
      corr_yes |>
      filter(V1 == ms_v) |>
      pull(V2) |>
      as.character()

    preds <-
      c(preds_from_v, preds_from_ms) |>
      unique()

    preds <-
      preds[
        preds %in% data_names &
        preds != v
      ]

    pred_matrix[v, preds] <- 1
  }

  pred_matrix
}
```

```{r Create MICE predictor matrix}
pred_matrix <-
  correlation_matrix$data |>
  make_mice_pred_matrix(colnames(select(train_data_outlier_rm, -id)))
```

```{r Fit and save MICE model, include=FALSE}
imp_results <-
  mice(
    data = select(train_data_outlier_rm, -id)
    , method = 'pmm'
    , m = 10
    , seed = seed
    , predictorMatrix = pred_matrix
    , print = FALSE
  )

saveRDS(imp_results, "data/imp_results.rds")
```

```{r Impute missing values for all set, include=FALSE}
proc_data_imputed <-
  imp_results |>
  mice.mids(newdata = select(proc_data_outlier_rm, -id))|>
  complete(1)

proc_data_imputed <-
  cbind(select(proc_data_outlier_rm, id), proc_data_imputed)
```

```{r Obtain train set from imputed dataset, include=FALSE}
train_data_imputed <- slice(proc_data_imputed, match(train_id, id))
```

# Regression analysis

```{r Scale of categorical predictors, include=FALSE}
cat_scale <-
  train_data_imputed |>
  select_if(is.factor) |>
  lapply(levels) |>
  imap(~ data.frame(predictor = .y, level = .x)) |>
  reduce(rbind)
```

```{r Show scale of categorical predictors, echo=FALSE}
show_table(cat_scale, "cat_scale", "Scale of categorical predictors")
```

```{r Obtain possible term in the regression results, include=FALSE}
term_data <-
  train_data_imputed |>
  imap(
    ~ data.frame(
        covariate = .y
        , term =
          c("(Intercept)"
            , list(levels, \(x) "")[[ifelse(is.factor(.x), 1, 2)]](.x)
          )
      )
  ) |>
  reduce(rbind) |>
  mutate(category = ifelse(term == "(Intercept)", "", term)) |>
  mutate(term = ifelse(term == "(Intercept)", term, paste0(covariate, term)))
```

```{r Obtain candidate predictors, include=FALSE}
predictor_candidates <-
  train_data_imputed |>
  select(-id, -outcome) |>
  colnames()
```

## Univariate regression analysis

```{r Determine formula for univariate regression analysis, include=FALSE}
unireg_formula <-
  predictor_candidates |>
  lapply(c, "outcome") |>
  lapply(rev) |>
  lapply(paste0, collapse = " ~ ")

unireg_formula <-
  unireg_formula |>
  `names<-`(unireg_formula) |>
  lapply(as.formula)
```

```{r Conduct univariate regression analysis, include=FALSE}
if(class(train_data_imputed$outcome) == "numeric"){
  unireg <-
    unireg_formula |>
    lapply(lm, data = train_data_imputed)
}else{
  unireg <-
    unireg_formula |>
    lapply(glm, family = binomial(), data = train_data_imputed)
}
```

```{r Create custom function to obtain covariate per formula}
obtain_covariate_per_formula <-function(data){
  data |>
    imap(
      ~ data.frame(
          formula = .y
          , outcome = as.character(.x)[2]
          , predictor = str_split_fixed(as.character(.x)[3], " \\+ ", 2)[, 1]
          , other = str_split_fixed(as.character(.x)[3], " \\+ ", 2)[, 2]
          , covariate = as.character(.x)[3]
        )
    ) |>
    reduce(rbind) |>
    separate_rows(covariate, sep = " \\+ ")
}
```

```{r Obtain covariate per formula for univariate regression, include=FALSE}
unireg_covariate_data <- obtain_covariate_per_formula(unireg_formula)
```

```{r Create custom function to obtain reg results with covariate and term}
obtain_reg_results <- function(data, covariate_data, term_data){
  covariate_data <-
    covariate_data |>
    left_join(
      term_data
      , by = join_by(covariate)
      , relationship = "many-to-many"
    )
  
  data |>
    lapply(tidy) |>
    imap(~ mutate(.x, formula = .y)) |>
    reduce(rbind) |>
    left_join(covariate_data, by = join_by(formula, term))  |>
    filter(
      (term == "(Intercept)" & covariate == predictor)
      | term != "(Intercept)"
    )
}
```

```{r Create custom function to compute estimate and 95% CI from regression}
compute_est_reg <- function(data, OR = TRUE){
  if(OR){
    results <-
      data |>
      mutate(
        OR = exp(estimate)
        , LB = exp(estimate - std.error)
        , UB = exp(estimate + std.error)
        , estimate = OR
      )
  }else{
    results <-
      data |>
      mutate(
        LB = estimate - std.error
        , UB = estimate + std.error
      )
  }
  
  results |>
    select(outcome, predictor, term, category, p.value, estimate, LB, UB)
}
```

```{r Compute OR and 95% CI for univariate regression analysis, include=FALSE}
if(class(train_data_imputed$outcome) == "numeric"){
  unireg_results <-
    unireg |>
    obtain_reg_results(unireg_covariate_data, term_data) |>
    compute_est_reg(OR = FALSE)
}else{
  unireg_results <-
    unireg |>
    obtain_reg_results(unireg_covariate_data, term_data) |>
    compute_est_reg()
}
```

```{r Show OR and 95% CI for univariate regression analysis, echo=FALSE}
unireg_results |>
  show_table(
  "unireg_results"
  , "Univariate regression."
  , digits = c(p.value = 3, estimate = 2, LB = 2, UB = 2)
)  
```

```{r Significant predictors by univariate regression analysis, include=FALSE}
predictor_candidates_unireg <-
  unireg_results |>
  filter(term == predictor) |>
  filter(p.value <= 0.05) |>
  filter(LB > 1 | UB < 1) |>
  pull(predictor) |>
  unique()
```

## Collinearity by correlation matrix

```{r Filter matrix if predictors are significant in uni reg, include=FALSE}
predictor_candidates_cm <-
  pred_matrix |>
  as.data.frame() |>
  rownames_to_column(var = "predictor") |>
  select(-outcome) |>
  slice(match(predictor_candidates_unireg, predictor))
```

```{r Identify sig cor between each candidate predictor to the others, include=FALSE}
predictor_candidates_cm_sig <-
  predictor_candidates_cm |>
  gather(other, correlation, -predictor) |>
  filter(correlation == 1)
```

```{r List others that correlate with each CP and outcome , include=FALSE}
predictor_candidates_stat_cf <-
  predictor_candidates_cm_sig |>
  group_by(predictor) |>
  summarise(values = list(other)) |>
  deframe() |>
  lapply(intersect, predictor_candidates_unireg)
```

## Multivariate regression analysis

```{r Determine formula for multivariate regression analysis, include=FALSE}
multireg_formula <-
  unireg_formula |>
  lapply(as.character) |>
  imap(~ c(.x, predictor_candidates_stat_cf[[.x[3]]])) |>
  imap(~ paste0("outcome ~ ", paste0(.x[-1:-2], collapse = " + ")))

multireg_formula <-
  multireg_formula |>
  `names<-`(multireg_formula) |>
  lapply(as.formula)
```

```{r Conduct multivariate regression analysis, include=FALSE}
if(class(train_data_imputed$outcome) == "numeric"){
  multireg <-
    multireg_formula |>
    lapply(lm, data = train_data_imputed)
}else{
  multireg <-
    multireg_formula |>
    lapply(glm, family = binomial(), data = train_data_imputed)
}
```

```{r Obtain covariate per formula for multivariate regression, include=FALSE}
multireg_covariate_data <- obtain_covariate_per_formula(multireg_formula)
```

```{r Compute OR and 95% CI for multivariate regression analysis, include=FALSE}
if(class(train_data_imputed$outcome) == "numeric"){
  multireg_results <-
    multireg |>
    obtain_reg_results(multireg_covariate_data, term_data) |>
    compute_est_reg(OR = FALSE)
}else{
  multireg_results <-
    multireg |>
    obtain_reg_results(multireg_covariate_data, term_data) |>
    compute_est_reg()
}
```

```{r Show OR and 95% CI for multivariate regression analysis, echo=FALSE}
multireg_results |>
  show_table(
  "multireg_results"
  , "Multivariate regression."
  , digits = c(p.value = 3, estimate = 2, LB = 2, UB = 2)
)
```

```{r Significant predictors by multivariate regression analysis, include=FALSE}
predictor_candidates_multireg <-
  multireg_results |>
  filter(term == predictor) |>
  filter(p.value <= 0.05) |>
  filter(LB > 1 | UB < 1) |>
  pull(predictor) |>
  unique()
```

## Structural causal modeling (SCM)

```{r Select some of existing formulas that has >1 covariates, include=FALSE}
formula_multi_covariates <-
  multireg_covariate_data |>
  group_by(formula) |>
  filter(length(covariate) >1) |>
  ungroup()
```

```{r Pair outcome with each covariate, include=FALSE}
mc_outcome_covariate <-
  formula_multi_covariates |>
  select(V1 = outcome, V2 = covariate) |>
  unique()
```

```{r Pair each candidate predictor with other covariate, include=FALSE}
mc_predictor_other <-
  formula_multi_covariates |>
  select(V1 = predictor, V2 = covariate) |>
  filter(V1 != V2) |>
  unique()
```

```{r Bind the selected pairs of variables, include=FALSE}
mc_pair_undirected <-
  rbind(mc_outcome_covariate, mc_predictor_other) |>
  mutate(from = NA)
```

```{r Write the bound-selected pairs to CSV, include=FALSE}
write_csv(mc_pair_undirected, "inst/extdata/mc_pair_undirected.csv")
```

```{r Determine causal direction per pair, include=FALSE}
mc_pair_directed <-
  read_csv("inst/extdata/mc_pair_directed.csv", show_col_types = FALSE)
```

```{r Obtain causal diagram data, include=FALSE}
causal_diagram_data <-
  mc_pair_directed |>
  mutate(to = ifelse(from == 1, 2, 1)) |>
  mutate(
    from = ifelse(from == 1, V1, V2)
    , to = ifelse(to == 1, V1, V2)
  ) |>
  select(from, to)
```

```{r Create custom function to plot causal diagram}
causal_diagram_plotter <- function(data, padding = 0.1, title = NULL){
  canvas_size <-
    data |>
    select(x, y, xend, yend) |>
    gather() |>
    summarize(min_value = min(value), max_value = max(value)) |>
    mutate(
      min_value = min_value - padding * (max_value - min_value)
      , max_value = max_value + padding * (max_value - min_value)
    ) |>
    gather() |>
    select(x = value, y = value)
  
  data |>
    ggplot(aes(x, y)) +
    geom_blank(data = canvas_size) +
    geom_nodes(aes(color = voi), size = 6) +
    geom_edges(
      aes(xend = xend, yend = yend, color = voi)
      , arrow = arrow(length = unit(6, "pt"), type = "closed"), curvature = 0.1
    ) +
    geom_label(
      aes(color = voi, label = name)
      , hjust = -0.2, vjust = 1.2, show.legend = FALSE
    ) +
    scale_color_discrete("Variable of\ninterest") +
    theme_blank() +
    ggtitle(title)
}
```

```{r Obtain causal diagram per formula, include=FALSE}
cd_formula <-
  formula_multi_covariates |>
  group_by(formula) |>
  summarise(values = list(covariate)) |>
  deframe() |>
  imap(
    ~ causal_diagram_data |>
      filter(from %in% c("outcome", .x) & to %in% c("outcome", .x))
  )

cd_graph <- lapply(cd_formula, graph_from_data_frame)

set.seed(seed)
cd_layout <- lapply(cd_graph, layout_as_tree, circular = TRUE, root = "outcome")

cd_plot_data <-
  cd_graph |>
  imap(~ ggnetwork(.x, cd_layout[[.y]])) |>
  imap(~ mutate(.x, pair = str_split_fixed(.y, " \\+ ", 2)[[1]])) |>
  lapply(separate, pair, c("outcome", "predictor"), sep = " ~ ") |>
  lapply(
    mutate
    , voi =
      case_when(
        name == outcome ~ "Outcome"
        , name == predictor ~ "Predictor"
        , TRUE ~ "Other"
      )
  )

cd_plot <-
  cd_plot_data |>
  imap(~ causal_diagram_plotter(.x, title = which(names(cd_plot_data) == .y)))

cd_plot <- ggarrange(plotlist = cd_plot, nrow = length(cd_plot))
```

```{r Create custom function to remove variables from formula}
remove_from_formula <- function(formula, variables){
  formula <- as.character(formula)
  
  formula <-
    formula |>
    lapply(str_split_1, " \\+ ") |>
    lapply(setdiff, variables) |>
    lapply(paste0, collapse = " + ") |>
    reduce(c)
  
  as.formula(paste0("outcome ~ ", formula[3]))
}
```

```{r CD >1 covariates, echo=FALSE, fig.height=560/25.4, fig.width=190/25.4}
cd_plot
```

```{r Caption CD >1 covariates, echo=FALSE}
show_figure_legend("cd_plot", "Causal diagram with >1 covariates.")
```

```{r Identify colliders, include=FALSE}
colliders <-read_csv("inst/extdata/colliders.csv", show_col_types = FALSE)
```

```{r Remove colliders from formula, include=FALSE}
removed_from_formula <- lapply(names(cd_plot_data), as.formula)

removed_from_formula <-
  colliders |>
  group_by(no) |>
  summarise(values = list(variable)) |>
  deframe() |>
  imap(~ remove_from_formula(removed_from_formula[[as.numeric(.y)]], .x)) |>
  `names<-`(names(cd_plot_data))
```

```{r Determine formula for multivariate regression analysis, include=FALSE}
scm_formula <- setdiff(names(multireg_formula), names(removed_from_formula))

scm_formula <-
  multireg_formula[scm_formula] |>
  c(removed_from_formula)

scm_formula <- scm_formula[names(multireg_formula)]

scm_formula <- `names<-`(scm_formula, scm_formula)
```

```{r Conduct SCM-multivariate regression analysis, include=FALSE}
if(class(train_data_imputed$outcome) == "numeric"){
  scm <-
    scm_formula |>
    lapply(lm, data = train_data_imputed)
}else{
  scm <-
    scm_formula |>
    lapply(glm, family = binomial(), data = train_data_imputed)
}
```

```{r Obtain covariate per formula for SCM-multivariate regression, include=FALSE}
scm_covariate_data <- obtain_covariate_per_formula(scm_formula)
```

```{r Compute OR and 95% CI for SCM-multivar regression analysis, include=FALSE}
if(class(train_data_imputed$outcome) == "numeric"){
  scm_results <-
    scm |>
    obtain_reg_results(scm_covariate_data, term_data) |>
    compute_est_reg(OR = FALSE)
}else{
  scm_results <-
    scm |>
    obtain_reg_results(scm_covariate_data, term_data) |>
    compute_est_reg()
}
```

```{r Show OR and 95% CI for SCM-multivar regression analysis, echo=FALSE}
scm_results |>
  show_table(
  "scm_results"
  , "Structural causal modeling using multivariate regression."
  , digits = c(p.value = 3, estimate = 2, LB = 2, UB = 2)
)
```

```{r Significant predictors by SCM-multivar regression analysis, include=FALSE}
predictor_candidates_scm <-
  scm_results |>
  filter(term == predictor) |>
  filter(p.value <= 0.05) |>
  filter(LB > 1 | UB < 1) |>
  pull(predictor) |>
  unique()
```

```{r Compared multivariate and SCM-multivariate regrssion results, include=FALSE}
compare_multireg_scm_results <-
  list(no = multireg_results, yes = scm_results) |>
  lapply(
    filter
    , predictor %in% c(predictor_candidates_multireg, predictor_candidates_scm)
  ) |>
  lapply(filter, term == predictor) |>
  imap(~ mutate(.x, SCM = .y)) |>
  reduce(rbind) |>
  arrange(outcome, predictor, term, category, SCM) |>
  select(outcome, predictor, term, category, SCM, everything())
```

```{r Show multivariate and SCM-multivariate regrssion results, echo=FALSE}
compare_multireg_scm_results |>
  show_table(
    "compare_multireg_scm_results"
    , "Compare multivariate regression without and with SCM."
    , digits = c(p.value = 3, estimate = 5, LB = 5, UB = 5)
  )
```

# Dimensional reduction

```{r Create custom function to numerize categorical variables-2, include=FALSE}
numerize_cat_var <- function(x){
 as.numeric(x) - 1 
}
```

```{r Select categorical variables, include=FALSE}
cat_data <- select_if(train_data, is.factor)
```

```{r Prepare cluster data-2, include=FALSE}
cluster_data <- train_data

if(ncol(cat_data) > 0){
  cluster_data <-
    cluster_data |>
    mutate_if(is.factor, numerize_cat_var) # Numerize categorical variables
}

cluster_data <-
  cluster_data |>
  column_to_rownames(var = "id") |>
  na.omit() # Remove rows with NA in any columns
```

## PCA

```{r Select variables for PCA}
# All variables except variable of interest
pca_var <- setdiff(colnames(cluster_data), "outcome")

# Selected variables (remove hash tag for commenting codes below)
# pca_var <- c("uta_ri", "uta_pi", "uta_pi", "sflt1", "plgf")
```

```{r Conduct PCA, include=FALSE}
pca_fit_var <-
  cluster_data |>
  select_at(pca_var) |>
  prcomp(scale. = TRUE) # Data standardization
```

```{r Create custom function to make scree plot}
scree_plotter <- function(pca){
  # Extract variance explained
  var_explained <-
    pca$sdev^2 / sum(pca$sdev^2)

  data <-
    data.frame(
      pc = seq_along(var_explained)
      , var = var_explained
    )

  ggplot(data, aes(x = pc, y = var)) +
    geom_line() +
    geom_point() +
    scale_x_continuous(breaks = data$pc) +
    xlab("Principal Component") +
    ylab("Proportion of Variance Explained") +
    theme(
      axis.title = element_text(size = 10)
    )
}
```

```{r Show scree plot, echo=FALSE, fig.height=75/25.4, fig.width=190/25.4}
scree_plotter(pca_fit_var)
```

```{r Caption scree plot, echo=FALSE}
show_figure_legend("scree_plotter(pca_fit_var)", "Scree plot.")
```

```{r Create custom function to make score plot with optional grouping}
pc_score_plotter <- function(
  pca
  , pc_x = 1
  , pc_y = 2
  , group = NULL
){
  # Extract scores
  scores <-
    as.data.frame(pca$x)

  # Prepare data
  data <-
    scores |>
    mutate(
      group = if(is.null(group)) "All" else as.factor(group)
    )

  # Variance explained for labels
  var_explained <-
    pca$sdev^2 / sum(pca$sdev^2)

  xlab <-
    paste0(
      "PC", pc_x, " ("
      , round(var_explained[pc_x] * 100, 1)
      , "%)"
    )

  ylab <-
    paste0(
      "PC", pc_y, " ("
      , round(var_explained[pc_y] * 100, 1)
      , "%)"
    )

  ggplot(data, aes(
    x = .data[[paste0("PC", pc_x)]]
    , y = .data[[paste0("PC", pc_y)]]
    , color = group
  )) +
    geom_point(size = 3) +
    xlab(xlab) +
    ylab(ylab) +
    theme(
      axis.title = element_text(size = 10)
      , legend.title = element_blank()
    )
}
```

```{r Show score plot, echo=FALSE, fig.height=108/25.4, fig.width=190/25.4}
pc_score_plotter(pca_fit_var)
```

```{r Caption score plot, echo=FALSE}
show_figure_legend("pc_score_plotter(pca_fit_var)", "Score plot.")
```

```{r Create custom function to make loading plot}
pc_loading_plotter <- function(
  pca
  , pc_x = 1
  , pc_y = 2
  , labels = TRUE
){
  # Extract loadings
  loadings <-
    as.data.frame(pca$rotation)

  loadings$variable <-
    rownames(loadings)

  # Variance explained
  var_explained <-
    pca$sdev^2 / sum(pca$sdev^2)

  xlab <-
    paste0(
      "PC", pc_x, " ("
      , round(var_explained[pc_x] * 100, 1)
      , "%)"
    )

  ylab <-
    paste0(
      "PC", pc_y, " ("
      , round(var_explained[pc_y] * 100, 1)
      , "%)"
    )

  p <-
    ggplot(loadings) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_vline(xintercept = 0, linetype = "dashed") +

    # Arrows
    geom_segment(
      aes(
        x = 0, y = 0
        , xend = .data[[paste0("PC", pc_x)]]
        , yend = .data[[paste0("PC", pc_y)]]
      )
      , arrow = arrow(length = unit(0.2, "cm"), type = "closed")
    ) +

    xlab(xlab) +
    ylab(ylab) +
    theme(
      axis.title = element_text(size = 10)
    )

  if(labels){
    p <-
      p +
      geom_text(
        aes(
          x = .data[[paste0("PC", pc_x)]]
          , y = .data[[paste0("PC", pc_y)]]
          , label = variable
        )
        , vjust = -0.5
        , size = 3
      )
  }

  return(p)
}
```

```{r Show loading plot, echo=FALSE, fig.height=190/25.4, fig.width=190/25.4}
pc_loading_plotter(pca_fit_var)
```

```{r Caption loading plot, echo=FALSE}
show_figure_legend("pc_loading_plotter(pca_fit_var)", "Loading plot.")
```

```{r Prepare PC1 and PC2 for clustering analysis, include=FALSE}
pc12_var_data <-
  pca_fit_var$x |>
  as.data.frame() |>
  select(PC1, PC2)
```

## Clustering analysis

```{r Choose between hierarhical and k-means clustering}
which_clustering <- c(`1` = "hierarchical", `2` = "k-means")[1]
```

```{r Choose clustering parameters}
# For hierarchical clustering
hclust_distance_method <-
  c(
    `1` = "gower"
    , `2` = "euclidean"
    , `3` = "binary"
    , `4` = "canberra"
    , `5` = "manhattan"
    , `6` = "maximum"
    , `7` = "minkowski"
  )[1]

hclust_agglomeration_method <-
  c(
    `1` = "average"
    , `2` = "ward.D2"
    , `3` = "complete"
    , `4` = "single"
  )[1]

h_cluster_n <- 4 # Change later after assessing the plot

# For k-means clustering
k_cluster <- NULL  # If 2 clusters are expected, change to k_cluster <- 2

kmeans_k_method <-
  c(`1` = "elbow"
    , `2` = "silhouette"
  )[1]

kmeans_k_range <- 2:10

silhouette_distance_method <-
  c(
    `1` = "gower"
    , `2` = "euclidean"
    , `3` = "binary"
    , `4` = "canberra"
    , `5` = "manhattan"
    , `6` = "maximum"
    , `7` = "minkowski"
  )[1]
```

```{r Create a function to compute obtain WCSS}
kmeans_wcss <- \(k, data){
  set.seed(seed)
  fit <- kmeans(x = data, centers = k)
  
  data.frame(k = k, metric = fit$tot.withinss)
}
```

```{r Create a function to compute obtain silhouette score}
kmeans_silhouette_score <- \(k, data){
  set.seed(seed)
  fit <- kmeans(x = data, centers = k)
  
  d <-
    if(silhouette_distance_method == "gower"){
      cluster::daisy(data, metric = "gower")
    }else{
      dist(data, method = silhouette_distance_method)
    }

  sil <- silhouette(fit$cluster, d)

  data.frame(k = k, metric = mean(sil[, "sil_width"]))
}
```

```{r Create a function to plot elbow or silhouette method}
kmeans_k_plotter <- function(
  data
  , k_range = 2:10
  , method = c("elbow", "silhouette")
){
  method <- match.arg(method)

  if(method == "elbow"){
    metric_data <- map_dfr(k_range, \(k) kmeans_wcss(k, data))

    ylab <- "Total Within-Cluster Sum of Squares"
  }else{
    metric_data <- map_dfr(k_range, \(k) kmeans_silhouette_score(k, data))

    ylab <- "Average Silhouette Width"
  }

  ggplot(metric_data, aes(x = k, y = metric)) +
    geom_line() +
    geom_point() +
    scale_x_continuous(breaks = k_range) +
    xlab("Number of Clusters (K)") +
    ylab(ylab) +
    theme(
      axis.title = element_text(size = 10)
    )
}
```

```{r Plot to choose k for k-means clustering if needed, include=FALSE}
if(which_clustering == "k-means" & is.null(k_cluster)){
  k_selection_plot <-
    kmeans_k_plotter(
      data = pc12_var_data
      , k_range = kmeans_k_range
      , method = kmeans_k_method
    )
}
```

```{r Show k plot, echo=FALSE, fig.height=90/25.4, fig.width=190/25.4}
if(which_clustering == "k-means" & is.null(k_cluster)){
  k_selection_plot
}
```

```{r Caption k plot, echo=FALSE}
if(which_clustering == "k-means" & is.null(k_cluster)){
  show_figure_legend("k_selection_plot", "k plot.")
}
```

```{r Conduct clustering analysis, include=FALSE}
if(which_clustering == "hierarchical"){
  if(hclust_distance_method == "gower"){
    clustering_fit <-
      pc12_var_data |>
      daisy(metric = "gower") |>
      hclust(method = hclust_agglomeration_method)
  }else{
    clustering_fit <-
      pc12_var_data |>
      dist(method = hclust_distance_method) |>
      hclust(method = hclust_agglomeration_method)
  }
}else if(which_clustering == "k-means"){
  set.seed(seed)
  clustering_fit <- kmeans(pc12_var_data, k_cluster)
}
```

```{r Obtain clusters, include=FALSE}
if(which_clustering == "hierarchical"){
  cluster_group <- cutree(clustering_fit, k = h_cluster_n)
}else{
  cluster_group <- clustering_fit$cluster
}
```

```{r Create custom function for PC-hclust plot}
pc_hclust_plotter <- function(hclust_fit, h_cluster){
  h_cluster <- factor(h_cluster)
  cols <- as.integer(h_cluster)

  plot(hclust_fit, cex = 0.6)

  labels <- hclust_fit$labels
  order <- hclust_fit$order
  usr <- par("usr")

  y1 <- usr[3] - 0.02 * diff(usr[3:4])
  y2 <- usr[3] - 0.06 * diff(usr[3:4])

  text(
    x = seq_along(order)
    , y = y2
    , labels = h_cluster[order]
    , srt = 90
    , adj = 1
    , xpd = NA
    , cex = 0.6
    , col = cols[order]
  )
}
```

```{r Show PC-hcl plot, echo=FALSE, fig.height=190/25.4, fig.width=190/25.4}
if(which_clustering == "hierarchical"){
  pc_hclust_plot <- pc_hclust_plotter(clustering_fit, cluster_group)
}
```

```{r Caption PC-hcl plot, echo=FALSE}
if(which_clustering == "hierarchical"){
  show_figure_legend("pc_hclust_plot", "PC-hclust plot.")
}
```

```{r Create cl-score plot, include=FALSE}
cl_score_plot <- pc_score_plotter(pca_fit_var, group = cluster_group)
```

```{r Show cl-score plot, echo=FALSE, fig.height=108/25.4, fig.width=190/25.4}
cl_score_plot
```

```{r Caption cl-score plot, echo=FALSE}
show_figure_legend("cl_score_plot", "Clustered score plot.")
```

## Ontology inference

```{r Transpose cluster data for ontology, include=FALSE}
ontology_data <-
  cluster_data |>
  select_at(pca_var) |>
  mutate_all(scale) |> # Data standardization
  t() |>
  as.data.frame()
```

```{r Conduct clustering analysis for ontology inference, include=FALSE}
if(hclust_distance_method == "gower"){
  ontology_fit <-
    ontology_data |>
    daisy(metric = "gower") |>
    hclust(method = hclust_agglomeration_method)
}else{
  ontology_fit <-
    ontology_data |>
    dist(method = hclust_distance_method) |>
    hclust(method = hclust_agglomeration_method)
}
```

```{r Choose height to group predictors }
ontology_height <- 0.18
```

```{r Obtain ontologies, include=FALSE}
ontology_group <- cutree(ontology_fit, h = ontology_height)
```

```{r Show ontoloy plot, echo=FALSE, fig.height=100/25.4, fig.width=190/25.4}
ontology_plot <- pc_hclust_plotter(ontology_fit, ontology_group)
```

```{r Caption ontology plot, echo=FALSE}
show_figure_legend("ontology_plot", "Ontology plot.")
```

```{r Create ontology group dataframe, include=FALSE}
ontology_group_df <-
  data.frame(ontology = sort(ontology_group)) |>
  rownames_to_column(var = "predictor") |>
  select(ontology, everything())
```

```{r Show ontology group dataframe, echo=FALSE}
show_table(ontology_group_df, "ontology_group_df", "Ontology group.")
```

## Ontology-wise PCA

```{r Determine predictors per ontology, include=FALSE}
ontology_predictors <-
  ontology_group_df |>
  group_split(ontology) |>
  lapply(pull, predictor)
```

```{r Conduct PCA per ontology, include=FALSE}
pca_fit_ontology <-
  ontology_predictors |>
  imap(~ select_at(cluster_data, .x)) |>
  lapply(prcomp, scale. = TRUE) # Data standardization
```

```{r Obtain PVE of PC1 from scree plot per ontology, include=FALSE}
ontology_pve <-
  pca_fit_ontology |>
  lapply(scree_plotter) |>
  imap(~ filter(.x$data, pc == 1)$var) |>
  unlist()

ontology_pve <-
  data.frame(ontology = seq(length(ontology_pve)), pve_pc1 = ontology_pve)
```

```{r Show PVE of PC1 from scree plot per ontology, echo=FALSE}
show_table(ontology_pve, "ontology_pve", "Ontology PVEs of PC1s.")
```

```{r}
combined_var <-
  ontology_group_df |>
  group_by(ontology) |>
  summarize(predictor_n = n()) |>
  filter(predictor_n > 1) |>
  left_join(ontology_pve, by = join_by(ontology)) |>
  filter(pve_pc1 >= 0.85) |>
  left_join(ontology_group_df, by = join_by(ontology)) |>
  group_split(ontology)

combined_var_ontology <-
  combined_var |>
  lapply(pull, ontology) |>
  sapply(unique)

combined_var <-
  combined_var |>
  lapply(pull, predictor) |>
  `names<-`(combined_var_ontology)

```

```{r Prepare data for all sets to combine varibles, include=FALSE}
comb_var_data <- proc_data_imputed

if(ncol(cat_data) > 0){
  comb_var_data <-
    comb_var_data |>
    mutate_if(is.factor, numerize_cat_var) # Numerize categorical variables
}

comb_var_data <-
  comb_var_data |>
  column_to_rownames(var = "id") |>
  na.omit() # Remove rows with NA in any columns
```

```{r Create custom function for combining variables to ontology}
combine_var_to_ontology <- function(data, ontology_id, var_comb){
  pca_fit_ontology[[as.numeric(ontology_id)]] |>
    predict(newdata = select_at(data, var_comb)) |>
    as.data.frame() |>
    select(PC1) |>
    `colnames<-`(paste("ontology", ontology_id))
}
```

```{r Combine variables for all set, include=FALSE}
proc_data_combined <-
  combined_var |>
  imap(~ combine_var_to_ontology(comb_var_data, .y, .x)) |>
  reduce(cbind)

proc_data_combined <-
  comb_var_data |>
  select_at(setdiff(colnames(comb_var_data), unlist(combined_var))) |>
  cbind(proc_data_combined) |>
  rownames_to_column(var = "id")
```

```{r Obtain train set from combined dataset, include=FALSE}
train_data_combined <- slice(proc_data_combined, match(train_id, id))
```







